---
title: "Hands-on_Ex04"
description: |
  I learn about the functions of spatstat package to perform spatial point patterns analysis.
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 09-04-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Loading the R packages

```{r echo=TRUE}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Spatial Data Wrangling

Importing the spatial data

```{r echo=TRUE}
childcare_sf <- st_read("../../../IS415/Hands-on_Ex/Hands-on_Ex04/data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

```{r echo=TRUE}
sg_sf <- st_read(dsn = "../../../IS415/Hands-on_Ex/Hands-on_Ex04/data", layer="CostalOutline")
```

```{r echo=TRUE}
mpsz_sf <- st_read(dsn = "../../../IS415/Hands-on_Ex/Hands-on_Ex04/data", layer = "MP14_SUBZONE_WEB_PL")
```

DIY: Ensure that they are projected in same projection system.

```{r echo=TRUE}
st_geometry(mpsz_sf)
st_geometry(sg_sf)
```

DIY: assign the correct crs to mpsz_sf and sg_sf simple feature data frames.

DIY: If necessary, changing the referencing system to Singapore national projected coordinate system.

```{r echo=TRUE}
mpszsf3414 <- st_set_crs(mpsz_sf, 3414)
sgsf3414 <- st_set_crs(sg_sf, 3414)
```

DIY: Using the mapping methods you learned in Hands-on Exercise 3, prepare a map as shown below.

```{r echo=TRUE}
tmap_mode("plot")
tm_shape(mpszsf3414) +
  tm_polygons() +
  qtm(childcare_sf, 
    fill = "DEPENDENCY")
```

Alternatively, a pin map

```{r echo=TRUE}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

Set mode to plot

```{r echo=TRUE}
tmap_mode('plot')
```


## Geospatial Data wrangling

Converting sf data frames to sp’s Spatial* class
DIY

```{r echo=TRUE}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

```{r}
displayfunction <- function(var) {
  var
}
```

Display all information of Spatial classes

```{r}
displayfunction(childcare)
displayfunction(mpsz)
displayfunction(sg)
```

Converting the Spatial* class into generic sp format

Note: spatstat requires the analytical data in ppp object form. There is no direct way to convert a Spatial* classes into ppp object. We need to convert the Spatial classes* into Spatial object first.

```{r echo=TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

Display sp objects 

```{r echo=TRUE}
childcare_sp
```

```{r echo=TRUE}
sg_sp
```

Converting the generic sp format into spatstat’s ppp format

```{r echo=TRUE}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

Plot childcare_ppp

```{r echo=TRUE}
plot(childcare_ppp)
```

Summary statistic of ppp object 

```{r echo=TRUE}
summary(childcare_ppp)
```

Handling duplicated points

check if any duplicates

```{r echo=TRUE}
any(duplicated(childcare_ppp))
```

Count the number of co-incidence point

```{r echo=TRUE}
multiplicity(childcare_ppp)
```

How many locations have more than one point event

```{r echo=TRUE}
sum(multiplicity(childcare_ppp) > 1)
```

View the locations of these duplicate point events

```{r echo=TRUE}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

```{r echo=TRUE}
tmap_mode('plot')
```

jittering to add a small perturbation to the duplicate points so that they do not occupy the exact same space

```{r echo=TRUE}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

DIY: check if any duplicates

```{r echo=TRUE}
any(duplicated(childcare_ppp_jit))
```

owin Object

owin is specially designed to represent polygonal region of singapore boundary

```{r echo=TRUE}
sg_owin <- as(sg_sp, "owin")
```

```{r echo=TRUE}
plot(sg_owin)
```

```{r echo=TRUE}
summary(sg_owin)
```

Combining point events object and owin object

Extract childcare events that are located within Singapore

```{r echo=TRUE}
childcareSG_ppp = childcare_ppp[sg_owin]
```


```{r echo=TRUE}
summary(childcareSG_ppp)
```

DIY: Plot childcareSG_ppp

```{r echo=TRUE}
plot(childcareSG_ppp)
```

## First-order Spatial Point Patterns Analysis

# Kernel Density Estimation

Computing kernel density estimation (Automatic bandwidth) 

```{r echo=TRUE}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

plot(kde_childcareSG_bw)
```

Retrieve the bandwidth used to compute the kde layer

```{r echo=TRUE}
bw <- bw.diggle(childcareSG_ppp)
bw
```

Rescalling KDE values

Convert the unit of measurement from meter to kilometer

```{r echo=TRUE}
childcareSG_ppp.km <- rescale(childcareSG_ppp, 1000, "km")
```

Rerun density with converted

```{r echo=TRUE}
kde_childcareSG.bw <- density(childcareSG_ppp.km, 
                              sigma=bw.diggle, 
                              edge=TRUE, 
                              kernel="gaussian")
plot(kde_childcareSG.bw)
```

Working with different automatic bandwidth methods

Functions to determine the bandwidth

```{r echo=TRUE}
bw.CvL(childcareSG_ppp.km)
```

```{r echo=TRUE}
bw.scott(childcareSG_ppp.km)
```

```{r echo=TRUE}
bw.ppl(childcareSG_ppp.km)
```

```{r echo=TRUE}
bw.diggle(childcareSG_ppp.km)
```

Comparison of output of using bw.diggle and bw.ppl methods

```{r echo=TRUE}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

Working with different kernel methods

```{r echo=TRUE}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

Fixed and Adaptive KDE

Computing KDE (Fixed Bandwidth)

```{r echo=TRUE}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```

Computing KDE (Adaptive Bandwidth)

```{r echo=TRUE}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```

Comparison of fixed and adaptive kernel density estimation outputs

```{r echo=TRUE}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```

Converting KDE output into grid object

```{r echo=TRUE}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```

Converting Gridded output into Raster

```{r echo=TRUE}
kde_childcareSG_bw_raster <- raster(gridded_kde_childcareSG_bw)
```

```{r echo=TRUE}
kde_childcareSG_bw_raster
```

CRS is NA. Assign Projection systems 

```{r echo=TRUE}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```

Visualising the output in tmap

Raster in cartographic quality map

```{r echo=TRUE}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```

Comparing Spatial Point Patterns using KDE of childcare of different areas

Extract study area

```{r echo=TRUE}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]
```

Plot target planning area

```{r echo=TRUE}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```

Convert Spatial point df into generic sp format

```{r echo=TRUE}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```

Create owin Object

```{r echo=TRUE}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```

Combine childcare point and study area

```{r echo=TRUE}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

Transform from m to km

```{r echo=TRUE}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")
```

Plot the four study areas

```{r echo=TRUE}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

Compute KDE (bw.diggle)

```{r echo=TRUE}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
```

Compute KDE (fixed bw)

```{r echo=TRUE}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

## Nearest Neighbour Analysis

Testing spatial point patterns using Clark and Evans Test

```{r echo=TRUE}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

Clark and Evans Test: Choa Chu Kang planning area

```{r echo=TRUE}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Clark and Evans Test: Tampines planning area

```{r echo=TRUE}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

## Second-order Spatial Point Patterns Analysis

# Analysing Spatial Point Process Using G-Function

CCK planning area

Computing G function

``` {r echo=TRUE}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```

complete Spatial Randomness Test

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

``` {r echo=TRUE}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```

``` {r echo=TRUE}
plot(G_CK.csr)
```

Tampines planning area

Computing G function

``` {r echo=TRUE}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

Complete Spatial Randomness Test

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

``` {r echo=TRUE}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```

``` {r echo=TRUE}
plot(G_tm.csr)
```

# Analysing Spatial Point Process Using F-Function

CCK Planning Area

Compute F-function

``` {r echo=TRUE}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

Complete Spatial Randomness Test

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

``` {r echo=TRUE}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```

```{r echo=TRUE}
plot(F_CK.csr)
```

Tampines Planning Area

Compute F-function

``` {r echo=TRUE}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

Complete Spatial Randomness Test

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

``` {r echo=TRUE}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```

```{r echo=TRUE}
plot(F_tm.csr)
```

# Analysing Spatial Point Process Using K-Function

CCK Planning Area

Compute K-Function estimate

```{r echo=TRUE}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

Complete Spatial Randomness Test

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

```{r echo=TRUE}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo=TRUE}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

Tampines Planning Area

Compute K-Function estimate

```{r echo=TRUE}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

Complete Spatial Randomness Test

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

```{r echo=TRUE}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo=TRUE}
plot(K_tm.csr, . - r ~ r, xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

# Analysing Spatial Point Process Using L-Function

CCK Planning Area

Computing L function

```{r echo=TRUE}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```


Performing the compelete Spatial Randomness Test

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value if smaller than alpha value of 0.001.

```{r echo=TRUE}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo=TRUE}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

Tampines planning area

```{r echo=TRUE}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

Performing the compelete Spatial Randomness Test

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value if smaller than alpha value of 0.001.

```{r echo=TRUE}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo=TRUE}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```


