---
title: "Hands-on_Ex02"
description: |
  I learn about importing, converting and projecting geospatial and aspatial data. 
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 08-30-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Getting Started

```{r echo=TRUE}
packages = c('sf', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

Importing polygon feature data in shapefile format

```{r echo=TRUE}
mpsz = st_read(dsn = "../../../IS415/Hands-on_Ex/Hands-on_Ex02/data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

Importing polyline feature data in shapefile form

```{r echo=TRUE}
cyclingpath = st_read(dsn = "../../../IS415/Hands-on_Ex/Hands-on_Ex02/data/geospatial", layer = "CyclingPath")
```
Importing GIS data in kml format

```{r echo=TRUE}
preschool = st_read("../../../IS415/Hands-on_Ex/Hands-on_Ex02/data/geospatial/pre-schools-location-kml.kml")
```

Working with st_geometry()

```{r echo=TRUE}
st_geometry(mpsz)
```

Working with glimpse()

```{r echo=TRUE}
glimpse(mpsz)
```

Working with head()

```{r echo=TRUE}
head(mpsz, n=5) 
```

Plotting the Geospatial Data

```{r echo=TRUE}
plot(mpsz)
```

Plotting the Geospatial Data only Geometry

```{r echo=TRUE}
plot(st_geometry(mpsz))
```

Plotting the Geospatial Data by using a specific attribute

```{r echo=TRUE}
plot(mpsz["PLN_AREA_N"])
```

Assigning EPSG code to a simple feature data frame (Coordinates Missing,  EPSG code for svy21 should be 3414)

```{r echo=TRUE}
st_crs(mpsz)
```

Assign the correct EPSG code to mpsz data frame

```{r echo=TRUE}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

check the CSR again (ESPG is 3414 now)

```{r echo=TRUE}
st_crs(mpsz3414)
```

Check geomtery of preschool

```{r echo=TRUE}
st_geometry(preschool)
```

Transforming the projection of preschool from wgs84 to svy21

```{r echo=TRUE}
preschool3414 <- st_transform(preschool, crs = 3414)
```

Check geometry of preschool3414

```{r echo=TRUE}
st_geometry(preschool3414)
```

Importing and Converting An Aspatial Data

```{r echo=TRUE}
listings <- read_csv("../../../IS415/Hands-on_Ex/Hands-on_Ex02/data/aspatial/listings.csv")
```

Glimpse listings data

```{r echo=TRUE}
glimpse(listings) 
```

Creating a simple feature data frame from an aspatial data frame

```{r echo=TRUE}
listings_sf <- st_as_sf(listings, coords = c("longitude", "latitude"), crs=4326) %>% st_transform(crs = 3414)
```

glimpse listings_sf

```{r echo=TRUE}
glimpse(listings_sf)
```

Geoprocessing with sf package (Buffering)

1. Compute the 5-meter buffers around cycling paths

```{r echo=TRUE}
buffer_cycling <- st_buffer(cyclingpath, dist=5, nQuadSegs = 30)
```

2. Calculating the area of the buffers

```{r echo=TRUE}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

3. sum() of Base R will be used to derive the total land involved

```{r echo=TRUE}
sum(buffer_cycling$AREA)
```

Geoprocessing with sf package (Point-in-polygon count)

1. identify pre-schools located inside each Planning Subzone by using st_intersects()
2. length() of Base R is used to calculate numbers of pre-schools that fall inside each planning subzone.

```{r echo=TRUE}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

summary statistics of the newly derived PreSch Count field by using summary() 

```{r echo=TRUE}
summary(mpsz3414$`PreSch Count`)
```

list the planning subzone with the most number of pre-school

```{r echo=TRUE}
top_n(mpsz3414, 1, `PreSch Count`)
```

Calculate the density of pre-school by planning subzone.

Uses st_area() of sf package to derive the area of each planning subzone.

```{r echo=TRUE}
mpsz3414$AREA <- st_area(mpsz3414)
```

Mutate() of dplyr package is used to compute the density

```{r echo=TRUE}
mpsz3414 <- mpsz3414 %>% mutate(`PreSch Density` = `PreSch Count`/`AREA` * 1000000)
```

Explorotary Data Analysis (EDA)

Histogram to reveal the distribution of PreSch Density

```{r echo=TRUE}
hist(mpsz3414$`PreSch Density`)
```

ggplot2 (histogram)

```{r echo=TRUE}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

ggplot2 (histogram)

```{r echo=TRUE}


ggplot(data=mpsz3414, aes(x= as.numeric(`PreSch Density`), y= `PreSch Count`))+ geom_point() + labs(x = "Pre-school density (per km sq)", y = "Pre-school count")
```


