---
title: "Hands-on_Ex06"
description: |
  I learnt how to derive spatial weights by using functions provided by spdep package and how to apply these spatial   weights to compute geographically referenced attributes.
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 09-18-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Install Packages
```{r echo=TRUE}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Data Wrangling

Import shapefile into r environment

```{r echo=TRUE}
hunan <- st_read(dsn = "data/geospatial", layer = "Hunan")
```

Import csv file into r environment

```{r echo=TRUE}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

Relational Join

This is performed by using left_join() of dplyr package.

```{r echo=TRUE}
hunan <- left_join(hunan,hunan2012)
```

Choropleth Map Distribution of GDPPC 2012

```{r echo=TRUE}
qtm(hunan, "GDPPC")
```

# Computing Contiguity Spatial Weights

Use poly2nb() of spdep package to compute contiguity weight matrices for the study area. The function builds a neighbours list based on regions with contiguous boundaries. The function takes in “queen” argument that takes TRUE or FALSE. FALSE will return list of first order neighbours using the Queen criteria.

## Computing (QUEEN) contiguity based neighbours

```{r echo=TRUE}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```
There are 88 area units in Hunan. The most connected area unit have 11 neighbours.

We can see the first for first polygon with the code below 

```{r echo=TRUE}
wm_q[[88]]
```

Retrieve the county name of Polygon ID=1 with the code chunk below

```{r echo=TRUE}
hunan$County[1]
```
Retrieve county names of the five neighboring polygons with the code chunk below

```{r echo=TRUE}
hunan$NAME_3[c(2,3,4,57,85)]
```

display the complete weight matrix by using str().

```{r echo=TRUE}
str(wm_q)
```

## Creating (ROOK) contiguity based neighbours

Compute Rook contiguity weight matrix.

```{r echo=TRUE}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

The result shows 88 area units. The most connected area unit have 10 neighbours.

Visualising contiguity weights

```{r echo=TRUE}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

```{r echo=TRUE}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

Put longitude and latitude into the same object.

```{r echo=TRUE}
coords <- cbind(longitude, latitude)
```

```{r echo=TRUE}
head(coords)
```

Plotting Queen contiguity based neighbours maps

```{r echo=TRUE}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```
Plotting Rook contiguity based neighbours maps

```{r echo=TRUE}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

Side by Side Comparison

```{r echo=TRUE}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")
```

## Computing distance based neighbours

derive distance-based weight matrices by using dnearneigh() of spdep package.

The function identifies neighbours of region points by Euclidean distance with a distance band with lower d1= and upper d2= bounds controlled by the bounds= argument. If unprojected coordinates are used and either specified in the coordinates object x or with x as a two column matrix and longlat=TRUE, great circle distances in km will be calculated assuming the WGS84 reference ellipsoid.

Determine the upper limit for distance band

```{r echo=TRUE}
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

Computing fixed distance weight matrix

```{r echo=TRUE}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

Display the content of wm_d62 weight matrix.

```{r echo=TRUE}
str(wm_d62)
```

Display the structure of the weight matrix with table() and card() of spdep.

```{r echo=TRUE}
table(hunan$County, card(wm_d62))
```

```{r echo=TRUE}
n_comp <- n.comp.nb(wm_d62)
n_comp$nc
```

```{r echo=TRUE}
table(n_comp$comp.id)
```

Plotting fixed distance weight matrix

```{r echo=TRUE}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```
Red lines show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.

Side by Side

```{r echo=TRUE}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")
```

Computing adaptive distance weight matrix

```{r echo=TRUE}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
```

```{r echo=TRUE}
str(knn6)
```
Plotting distance based neighbours

```{r echo=TRUE}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## Weights based on IDW

derive a spatial weight matrix based on Inversed Distance method.

Compute the distances between areas by using nbdists() of spdep.

```{r echo=TRUE}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## Row-standardised weights matrix

Assign weights to each neighboring polygon

```{r echo=TRUE}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

Weight of the first polygon’s four neighbors type:

```{r echo=TRUE}
rswm_q$weights[10]
```

derive a row standardised distance weight matrix with the same method 

```{r echo=TRUE}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids
```

```{r echo=TRUE}
rswm_ids$weights[1]
```

```{r echo=TRUE}
summary(unlist(rswm_ids$weights))
```

# Application of Spatial Weight Matrix

## Spatial lag with row-standardized weights

Compute the average neighbor GDPPC value for each polygon

```{r echo=TRUE}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

Append the spatially lag GDPPC values onto hunan SpatialPolygonDataFrame

```{r echo=TRUE}
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)
```

```{r echo=TRUE}
head(hunan)
```

GDPPC and spatial lag GDPPC

```{r echo=TRUE}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)

```

Spatial window sum

Assign knn6 to a new variable

```{r echo=TRUE}
knn6a <- knn6
```

add the diagonal element to the neighbour list

```{r echo=TRUE}
include.self(knn6a)
```

assign binary weights to the neighbour structure

```{r echo=TRUE}
binary.knn6 <- lapply(knn6a, function(x) 0*x+1)
binary.knn6[1]
```
nb2listw() and glist() to explicitly assign weight values.

```{r echo=TRUE}
wm_knn6 <- nb2listw(knn6a, glist = binary.knn6, style = "B")
```

compute the lag variable with lag.listw()

```{r echo=TRUE}
lag_knn6 <- lag.listw(wm_knn6, hunan$GDPPC)
```

convert the lag variable listw object into a data.frame

```{r echo=TRUE}
lag.list.knn6 <- list(hunan$NAME_3, lag.listw(wm_knn6, hunan$GDPPC))
lag_knn6.res <- as.data.frame(lag.list.knn6)
colnames(lag_knn6.res) <- c("NAME_3", "lag_sum GDPPC")
```

append lag_sum GDPPC values to hunan sf data.frame

```{r echo=TRUE}
hunan <- left_join(hunan, lag_knn6.res)
```

Plot GDPPC and lag_sum GDPPC

```{r echo=TRUE}
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```

