---
title: "In-class Exercise 4"
description: |
  A short description of the post.
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 09-06-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Installing and Loading the R packages

- sf is mainly used for importing geospatial data
- tidyverse to read text file and data manupulation
- tmap to do mapping
- spatstat to perform spatial point method analysis
- maptools to convert spatial data from spatial object class into ppp format
- raster to handle raster-based geospatial data

```{r echo=TRUE}
packages <- c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

## Importing Geospatial Data

```{r echo=TRUE}
sg_sf <- st_read(dsn = "../../../IS415/In-class_Ex/In-class_Ex04/data/shapefile", layer="CostalOutline")
mpsz_sf <- st_read(dsn = "../../../IS415/In-class_Ex/In-class_Ex04/data/shapefile", layer="MP14_SUBZONE_WEB_PL")
```

## Importing Aspatial Data

```{r echo=TRUE}
childcare <- read_rds("../../../IS415/In-class_Ex/In-class_Ex04/data/rds/childcare.rds")
CHAS <- read_rds("../../../IS415/In-class_Ex/In-class_Ex04/data/rds/CHAS.rds")
```

### Converting from aspatial to geospatial

```{r echo=TRUE}
CHAS_sf <- st_as_sf(CHAS,
                    coords = c("X_COORDINATE","Y_COORDINATE"),
                      crs=3414)
```

note that childcare data lat lng is decimal.
Hence need to st_as_sf to crs 4326 before transforming to crs 3414

```{r echo=TRUE}
childcare$Lat <- as.numeric(childcare$Lat)
childcare$Lng <- as.numeric(childcare$Lng)
childcare_sf <- st_as_sf(childcare,
                    coords = c("Lng","Lat"),
                      crs=4326) %>%
  st_transform(crs=3414)
```

## Geospatial Data Wrangling

### Converting from sf to Spatial* classes

Notice that data will be loss along the way...

```{r echo=TRUE}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

### Converting Spatial* classes to sp format
```{r echo=TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

### Converting from sf to Spatial* classes

```{r echo=TRUE}
chilcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```

```{r echo=TRUE}

```
