---
title: "In-class Exercise 3"
description: |
  In this hands-on exercise, I learn about percentile, box, rate and relative risk map.
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 08-30-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Getting Started

- This code chunk create a list of R packages needed and install them if they are not installed.
```{r echo=TRUE}
packages = c('sf', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Cleaning RDS file

- This coded chuck reads the RDS file from the path and display the data.
```{r echo=TRUE}
mpsz_pop2020 <- read_rds("../../../IS415/Hands-on_Ex/Hands-on_Ex03/data/rds/mpszpop2020.rds")

knitr::kable(head(mpsz_pop2020))
```

- Boxplot
```{r echo=TRUE}
ggplot(data=mpsz_pop2020, 
       aes(x = "", 
           y = AGED)) +
  geom_boxplot()
```

## Visualization

- Percentile Map

Exclude records with NA
```{r echo=TRUE}
mpsz_pop2020a <- mpsz_pop2020 %>%
  drop_na()

knitr::kable(head(mpsz_pop2020a))
```

Percentile
```{r echo=TRUE}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- mpsz_pop2020a["DEPENDENCY"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

Creating get variable function
```{r echo=TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>%
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

Plotting percentile map with tmap functions (hardcode)
```{r echo=TRUE}
percent <- c(0,.01,.1,.5,.9,.99,1)
var = get.var("DEPENDENCY", mpsz_pop2020a)
bperc <- quantile(var,percent)
tm_shape(mpsz_pop2020) + 
  tm_polygons() +
tm_shape(mpsz_pop2020a) +
  tm_fill("DEPENDENCY",
          title="DEPENDENCY",
          breaks=bperc,
          palette="Blues",
          labels=c("< 1%", "1% - 10%",
                   "10% - 50%",
                   "50% - 90%",
                   "90% - 99%",
                   "> 99%")) + 
  tm_borders() +
  tm_layout(title = "Percentile Map", title.position = c("right", "bottom"))
```

Plotting percentile map with Percentile Mapping Function
```{r echo=TRUE}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var = get.var(vnam, df)
  bperc <- quantile(var,percent)
  tm_shape(df) + 
  tm_polygons() +
  tm_shape(df) +
    tm_fill(vnam,
            title=legtitle,
            breaks=bperc,
            palette="Blues",
            labels=c("< 1%", "1% - 10%",
                     "10% - 50%",
                     "50% - 90%",
                     "90% - 99%",
                     "> 99%")) + 
    tm_borders() +
    tm_layout(title = mtitle, title.position = c("right", "bottom"))
}
```

Testing Percentile Mapping Function (young)
```{r echo=TRUE}
youngmap <- percentmap("YOUNG", mpsz_pop2020a)
youngmap
```

Testing Percentile Mapping Function (aged)
```{r echo=TRUE}
agedmap <- percentmap("YOUNG", mpsz_pop2020a)
agedmap
```

- Boxmap

Boxbreaks Function
```{r echo=TRUE}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {# no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) {# no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```


Creating get variable function
```{r echo=TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>%
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

Testing the Boxbreaks Function on AGED using mpsz_pop2020 will result in error.

Since missing values and NaN's are not allowed, we will exclude AGED = NA

```{r echo=TRUE}
mpsz_pop2020b <- mpsz_pop2020 %>%
  filter(AGED>=0)
var <- get.var("AGED", mpsz_pop2020b)
boxbreaks(var)
```

Boxmap Function
```{r echo=TRUE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```

Testing Boxmap function (AGED)

```{r echo=TRUE}
boxmap("AGED", mpsz_pop2020b)
```

Boxmap Function (Missing Area Fix)
```{r echo=TRUE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(mpsz_pop2020) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```

Testing Boxmap function (AGED)

```{r echo=TRUE}
boxmap("AGED", mpsz_pop2020b)
```

- Raw Rate Map

Compute raw rate
```{r echo=TRUE}
mpsz_pop2020c <- mpsz_pop2020 %>%
  mutate(`AGED%` = (`AGED`/`TOTAL`)*100) %>%
  filter(`AGED%` >= 0)
```

Plot Raw rate map using Boxmap function 
```{r echo=TRUE}
var <- get.var("AGED%", mpsz_pop2020c)
boxbreaks(var)
boxmap("AGED%",mpsz_pop2020c)
```

- Relative Risk Map

```{r echo=TRUE}

```

```{r echo=TRUE}

```
