---
title: "Hands-on_Ex05"
description: |
  Network constrained Spatial Point Patterns Analysis (NetSPAA)
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 09-19-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Installing packages

```{r echo=TRUE}
packages = c('sp', 'rgdal', 'spNetwork', 'tmap')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Data

Import 

```{r echo=TRUE}
network <- readOGR(dsn="data/geospatial", 
                   layer="Punggol_St",
                   verbose = FALSE)
childcare <- readOGR(dsn="data/geospatial",
                     layer="Punggol_CC",
                     verbose = FALSE)
```

Print the content of network SpatialLineDataFrame and childcare SpatialPointsDataFrame using the code chunk below.

```{r echo=TRUE}
str(network)
str(childcare)
```

spNetwork is expecting the geospatial data contains complete CRS information.

In the code chunk below, spTransform() of sp package is used to assign EPSG code to the SpatialDataFrames. The epsg:3414 is the code for svy21.

```{r echo=TRUE}
childcare <-spTransform(childcare, CRS("+init=epsg:3414"))
network <- spTransform(network,CRS("+init=epsg:3414"))
```

# Visualising the Geospatial Data

```{r echo=TRUE}
plot(network)
plot(childcare,add=T,col='red',pch = 19)
```

Interactive manner

```{r echo=TRUE}
tmap_mode('view')
tm_shape(childcare)+
  tm_dots() +
tm_shape(network)+
  tm_lines()
```

# Network Constrained KDE (NetKDE) Analysis

## Preparing the lixels objects

Before computing NetKDE, the SpatialLines object need to be cut into lixels with a specified minimal distance

The length of a lixel, lx_length is set to 700m, and
The minimum length of a lixel, mindist is set to 350m.

```{r echo=TRUE}
lixels <- lixelize_lines(network,700,mindist = 350)
```

## Generating line centre points

Lines_center() of spNetwork will be used to generate a SpatialPointsDataFrame (i.e. samples) with line centre points as shown in the code chunk below.
 
```{r echo=TRUE}
samples <- lines_center(lixels)
```

## Performing NetKDE

compute the NetKDE

```{r echo=TRUE}
densities <- nkde(network, 
                  events = childcare,
                  w = rep(1,nrow(childcare)),
                  samples = samples,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

## Visualising NetKDE

Insert the computed density values (i.e. densities) into samples and lixels objects as density field.

```{r echo=TRUE}
samples$density <- densities
lixels$density <- densities
```

Rescale the density values from number of events per meter to number of events per kilometer.

```{r echo=TRUE}
# rescaling to help the mapping
samples$density <- samples$density*1000
lixels$density <- lixels$density*1000
```

```{r echo=TRUE}
tmap_mode('view')
tm_shape(lixels)+
  tm_lines(col="density")+
tm_shape(childcare)+
  tm_dots()
```

```{r echo=TRUE}
tmap_mode('plot')
```

# Network Constrained G- and K-Function Analysis

Ho: The observed spatial point events (i.e distribution of childcare centres) are uniformly distributed over a street network in Punggol Planning Area.

The CSR test is based on the assumption of the binomial point process which implies the hypothesis that the childcare centres are randomly and independently distributed over the street network.

If this hypothesis is rejected, we may infer that the distribution of childcare centres are spatially interacting and dependent on each other; as a result, they may form nonrandom patterns.

```{r echo=TRUE}
fun_childcare <- kfunctions(network, 
                             childcare,
                             start = 0, 
                             end = 1000, 
                             step = 50, 
                             width = 50, 
                             nsim = 50, 
                             resolution = 50,
                             verbose = FALSE, 
                             conf_int = 0.05)
```

visualise the ggplot2 object of k-function

```{r echo=TRUE}
fun_childcare$plotk
```

The blue line is the empirical network K-function of the childcare centres in Punggol planning area. The gray envelop represents the results of the 50 simulations in the interval 2.5% - 97.5%. Because the blue line between the distance of 250m-400m are below the gray area, we can infer that the childcare centres in Punggol planning area resemble regular pattern at the distance of 250m-400m.

```{r echo=TRUE}

```


