---
title: "Hands-on_Ex07"
description: |
  i have learn how to compute Global and Local Measures of Spatial Autocorrelation using spdep package
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 09-26-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Install Packages

- sf is use for importing and handling geospatial data in R,

- tidyverse is mainly use for wrangling attribute data in R,

- spdep will be used to compute spatial weights, global and local spatial autocorrelation statistics, and

- tmap will be used to prepare cartographic quality chropleth map.

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Data

## Shapefile

The code chunk below uses st_read() of sf package to import Hunan shapefile into R. The imported shapefile will be simple features Object of sf.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

## CSV

import Hunan_2012.csv into R by using read_csv() of readr package. The output is R data frame class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

## Relational Join

The code chunk below will be used to update the attribute table of hunan’s SpatialPolygonsDataFrame with the attribute fields of hunan2012 dataframe. This is performed by using left_join() of dplyr package.

```{r}
hunan <- left_join(hunan,hunan2012)
```
## Visualising Regional Development Indicator

The following code chunk prepare a basemap and a choropleth map showing the distribution of GDPPC 2012 by using qtm() of tmap package

```{r fig.width=8, fig.height=8}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

# Global Spatial Autocorrelation

## Computing Contiguity Spatial Weights

In the code chunk below, poly2nb() of spdep package is used to compute contiguity weight matrices for the study area, specifically compute Queen contiguity weight matrix.

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

88 neighbours, most connected area unit has 11 neighbours. Two area units with one neighbours

### Row-standardised weights matrix

assign weights to each neighboring polygon. The zero.policy=TRUE option allows for lists of non-neighbors

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

# Global Spatial Autocorrelation: Moran’s I

## Maron’s I test

The code chunk below performs Moran’s I statistical testing using moran.test() of spdep.

```{r}
moran.test(hunan$GDPPC, listw=rswm_q, zero.policy = TRUE, na.action=na.omit)
```

## Computing Monte Carlo Moran’s I

The code chunk below performs permutation test for Moran’s I statistic by using moran.mc() of spdep. A total of 1000 simulation will be performed.

```{r}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC, listw=rswm_q, nsim=999, zero.policy = TRUE, na.action=na.omit)
bperm
```

## Visualising Monte Carlo Moran’s I

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Moran's I")
abline(v=0, col="red") 
```

# Global Spatial Autocorrelation: Geary’s

## Geary’s C test

The code chunk below performs Geary’s C test for spatial autocorrelation by using geary.test() of spdep.

```{r}
geary.test(hunan$GDPPC, listw=rswm_q)
```

## Computing Monte Carlo Geary’s C

The code chunk below performs permutation test for Geary’s C statistic by using geary.mc() of spdep.

```{r}
set.seed(1234)
bperm=geary.mc(hunan$GDPPC, listw=rswm_q, nsim=999)
bperm
```

## Visualising the Monte Carlo Geary’s C

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Geary c")
abline(v=1, col="red") 
```

# Spatial Correlogram

examine patterns of spatial autocorrelation in your data or model residuals

## Compute Moran’s I correlogram

sp.correlogram() of spdep package is used to compute a 6-lag spatial correlogram of GDPPC.
 
```{r}
MI_corr <- sp.correlogram(wm_q, hunan$GDPPC, order=6, method="I", style="B")
plot(MI_corr)
```

## Compute Geary’s C correlogram and plot

p.correlogram() of spdep package is used to compute a 6-lag spatial correlogram of GDPPC

```{r}
GC_corr <- sp.correlogram(wm_q, hunan$GDPPC, order=6, method="C", style="W")
plot(GC_corr)
```

# Cluster and Outlier Analysis

## Computing local Moran’s I

To compute local Moran’s I, the localmoran() function of spdep will be used. It computes Ii values, given a set of zi values and a listw object providing neighbour weighting information for the polygon associated with the zi values.

```{r}
fips <- order(hunan$County)
localMI <- localmoran(hunan$GDPPC, rswm_q)
head(localMI)
```

The code chunk below list the content of the local Moran matrix derived by using printCoefmat().

```{r}
printCoefmat(data.frame(localMI[fips,], row.names=hunan$County[fips]), check.names=FALSE)
```

## Mapping the local Moran’s I

Append the local Moran’s I dataframe (i.e. localMI) onto hunan SpatialPolygonDataFrame
 
```{r}
hunan.localMI <- cbind(hunan,localMI) %>%
  rename(Pr.Ii = Pr.z....E.Ii..)
```

### Mapping local Moran’s I values

```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty",
          palette = "RdBu",
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)
```

### Mapping local Moran’s I p-values

The code chunks below produce a choropleth map of Moran’s I p-values by using functions of tmap package.

```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)
```

### Mapping both local Moran’s I values and p-values

```{r}
localMI.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty", 
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)

pvalue.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)

tmap_arrange(localMI.map, pvalue.map, asp=1, ncol=2)
```

# Creating a LISA Cluster Map

## Plotting Moran scatterplot

The code chunk below plots the Moran scatterplot of GDPPC 2012 by using moran.plot() of spdep.

```{r}
nci <- moran.plot(hunan$GDPPC, rswm_q,
                  labels=as.character(hunan$County), 
                  xlab="GDPPC 2012", 
                  ylab="Spatially Lag GDPPC 2012")
```

## Plotting Moran scatterplot with standardised variable

scale() to centers and scales the variable. The as.vector() added to the end is to make sure that the data type we get out of this is a vector, that map neatly into out dataframe.

```{r}
hunan$Z.GDPPC <- scale(hunan$GDPPC) %>% as.vector 
```

plot the Moran scatterplot using the code chunk below.

```{r}
nci2 <- moran.plot(hunan$Z.GDPPC, rswm_q,
                   labels=as.character(hunan$County),
                   xlab="z-GDPPC 2012", 
                   ylab="Spatially Lag z-GDPPC 2012")
```

## Preparing LISA map classes

```{r}
quadrant <- vector(mode="numeric",length=nrow(localMI))
```

centers the variable of interest around its mean.

```{r}
DV <- hunan$GDPPC - mean(hunan$GDPPC)
```

centering the local Moran’s around the mean.

```{r}
C_mI <- localMI[,1] - mean(localMI[,1]) 
```

set a statistical significance level for the local Moran.

```{r}
signif <- 0.05 
```

define the high-high, low-low, low-high and high-low categories.

```{r}
quadrant[DV >0 & C_mI>0] <- 4      
quadrant[DV <0 & C_mI<0] <- 1      
quadrant[DV <0 & C_mI>0] <- 2
quadrant[DV >0 & C_mI<0] <- 3
```

places non-significant Moran in the category 0.

```{r}
quadrant[localMI[,5]>signif] <- 0
```

## Plotting LISA map

```{r}
hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)
```

plot both the local Moran’s I values map and its corresponding p-values map next to each other

```{r}
gdppc <- qtm(hunan, "GDPPC")

hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

LISAmap <- tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)

tmap_arrange(gdppc, LISAmap, asp=1, ncol=2)
```

# Hot Spot and Cold Spot Area Analysis

## Hot Spot and Cold Spot Area Analysis

### Deriving the centroid

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

```{r}
coords <- cbind(longitude, latitude)
```

### Determine the cut-off distance

```{r}
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

### Computing fixed distance weight matrix

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

```{r}
wm62_lw <- nb2listw(wm_d62, style = 'B')
summary(wm62_lw)
```

### Computing adaptive distance weight matrix

```{r}
knn <- knn2nb(knearneigh(coords, k=8))
knn
```

```{r}
knn_lw <- nb2listw(knn, style = 'B')
summary(knn_lw)
```

# Computing Gi statistics

## Gi statistics using fixed distance

```{r}
fips <- order(hunan$County)
gi.fixed <- localG(hunan$GDPPC, wm62_lw)
gi.fixed
```

Join the Gi values to their corresponding hunan sf data frame

```{r}
hunan.gi <- cbind(hunan, as.matrix(gi.fixed)) %>%
  rename(gstat_fixed = as.matrix.gi.fixed.)
```


## Mapping Gi values with fixed distance weights

```{r}
gdppc <- qtm(hunan, "GDPPC")

Gimap <-tm_shape(hunan.gi) +
  tm_fill(col = "gstat_fixed", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)
```

## Gi statistics using adaptive distance

```{r}
fips <- order(hunan$County)
gi.adaptive <- localG(hunan$GDPPC, knn_lw)
hunan.gi <- cbind(hunan, as.matrix(gi.adaptive)) %>%
  rename(gstat_adaptive = as.matrix.gi.adaptive.)
```

## Mapping Gi values with adaptive distance weights

```{r}
gdppc <- qtm(hunan, "GDPPC")

Gimap <- tm_shape(hunan.gi) +
  tm_fill(col = "gstat_adaptive", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)
```