---
title: "Take Home Exercise 02"
description: |
  Analysing and Visualising the distribution of Airbnb listings in 2019 and 2021, and how Covid-19 affected the distribution.
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 09-16-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    number_sections: true
    code_folding: true
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1 Introduction

In recent years, the emergence of numerous peer-based business models has empowered individuals across the globe to become micro-entrepreneurs, earning money from their idle property and spare time. This phenomenon, entitled ‘the sharing economy’, has seen unprecedented growth in terms of user numbers, enabling new avenues of economic and social interaction (Sundararajan, 2016).

Since its inception in 2008, Airbnb has expanded into over 34,000 cities across 191 countries. Uncommonly for a sharing economy company, Airbnb transitioned into profitability in 2016, demonstrating proof-of-concept for the validity of sharing economy within the global market (Stone & Zaleski, 2017).

Interestingly, Singapore is one of the global city that has yet to legalise short-term rentals offered by platforms such as Airbnb (Read more at https://www.todayonline.com/singapore/short-term-home-sharing-remains-illegal-singapore-airbnb-disappointed). However, during my recent visit to Inside Airbnb (http://insideairbnb.com/about.html), an independent, non-commercial set of tools and data that allows anyone to explore how Airbnb is really being used in cities around the world, it came to my attention that there are data sets for Singapore.

This analysis consist of two sections:
1) Airbnb Distribution in 2019
2) Impact of COVID-19

# 2 Data

- Airbnb Listings (June 2019 & July 2021) is the Airbnb Listing in Singapore for the year 2019 and 2021. It is in csv format, and can be taken from http://insideairbnb.com/get-the-data.html
- Hotels data  is the list of hotels in Singapore. It is in csv format, and can be extracted from SLA OneMap Service using onemapsapi.
- Tourist Attractions data is the list of tourist attractions in Singapore. It is in csv format, and can be extracted from SLA OneMap Service using onemapsapi.
- MrtLrtStnPtt contains all the Mrt and Lrt stations in Singapore. It is in ESRI shapefile format, and can be taken from https://datamall.lta.gov.sg/
- MP14_SUBZONE_WEB_PL is URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format, and can be taken from https://www.data.gov.sg/.
- CostalOutline is the national boundary of Singapore. It is in ESRI shapefile format, and can taken from SLA website.

# 3 Setting Up the Environment

The following code chunk install the following packages:

- maptools to convert Spatial objects into ppp format
- sf to handle geospatial data
- raster to convert image output generate by spatstat into raster format.
- spatstat to perform 1st and 2nd-order spatial point patterns analysis and derive kernel density estimation (KDE) layer
- tmap to create choropleth maps
- tidyr to tidy messy data
- readr to import csv

```{r}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap', 'tidyr', 'readr')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```


# 4 Spatial Data Wrangling

## 4.1 Importing Geospatial Data 

### 4.1.1 MrtLrt

#### 4.1.1.1 Importing

The following code uses st_read function of sf package to import MrtLrt geospatial data sets.

```{r}
mrtlrt_sf <- st_read(dsn = "data/geospatial", layer="MRTLRTStnPtt")
```

The MrtLrt data have 185 features with 3 fields. The projected CRS is SVY21. The geometry type is in point. 

#### 4.1.1.2 Check CRS

Next, we use st_crs function of sf to check the CRS.

```{r}
st_crs(mrtlrt_sf)
```

From the result above, even though the mrtlrt_sf is projected in svy21, the last line indicates that the EPSG is 9001. It is wrong because the correct EPSG code for svy21 should be 3414.

#### 4.1.1.3 Assign EPSG code to a simple feature data frame

The following code chunk will assign the correct EPSG code to mrtlrt_sf using st_set_crs function of sf package

```{r}
mrtlrt_sf <- st_set_crs(mrtlrt_sf, 3414)
```

Now that we have assign the correct EPSG code to mrtlrt_sf, we will check the CRS again using st_crs function of sf

#### 4.1.1.4 Check CRS again

```{r}
st_crs(mrtlrt_sf)
```

From the result above, we have successfully assign EPSG 3414 to mrtlrt_sf.

### 4.1.2 CostalOutline

#### 4.1.2.1 Importing

The following code uses st_read function of sf package to import CostalOutline geospatial data sets.

```{r}
sg_sf <- st_read(dsn = "data/geospatial", layer="CostalOutline")
```

The CostalOutline data have 60 features with 4 fields. The projected CRS is SVY21. The geometry type is polygon.

#### 4.1.2.2 Check CRS

```{r}
st_crs(sg_sf)
```

Similarly, even though the sg_sf is projected in svy21, the last line indicates that the EPSG is 9001. It is wrong because the correct EPSG code for svy21 should be 3414.

#### 4.1.2.3 Assign EPSG code to a simple feature data frame

The following code chunk will assign the correct EPSG code to sg_sf using st_set_crs function of sf package

```{r}
sg_sf <- st_set_crs(sg_sf, 3414)
```

Now that we have assign the correct EPSG code to sg_sf, we will check the CRS again using st_crs function of sf

#### 4.1.2.4 Check CRS again

```{r}
st_crs(sg_sf)
```
We have successfully set the EPSG to 3414 for sg_sf.

### 4.1.3 MPSZ

#### 4.1.3.1 Importing
The following code uses st_read function of sf package to import MP14_SUBZONE_WEB_PL geospatial data sets.

```{r}
mpsz_sf <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

The mpsz data have 323 features with 15 fields. The projected CRS is SVY21. The geometry type is multipolygon.

#### 4.1.3.2 Check CRS

```{r}
st_crs(mpsz_sf)
```

Similarly, even though the mpsz_sf is projected in svy21, the last line indicates that the EPSG is 9001. It is wrong because the correct EPSG code for svy21 should be 3414.

#### 4.1.3.3 Assign EPSG code to a simple feature data frame

The following code chunk will assign the correct EPSG code to mpsz_sf using st_set_crs function of sf package

```{r}
mpsz_sf <- st_set_crs(mpsz_sf, 3414)
```

Now that we have assign the correct EPSG code to mpsz_sf, we will check the CRS again using st_crs function of sf

#### 4.1.3.4 Check CRS again

```{r}
st_crs(mpsz_sf)
```

We have successfully assign ESPG 3414 to mpsz_sf.

## 4.2 Importing Aspatial Data 

### 4.2.1 Hotel

#### 4.2.1.1 Importing

The following code uses read_csv function of readr package to import Hotel aspatial data sets.

```{r}
hotels <- read_csv("data/aspatial/hotels.csv")
```

#### 4.2.1.2 Examine the file

The following code uses list function of base R to examine the imported file

```{r}
list(hotels)
```

The data contains 422 rows and 9 columns. There are 'Lat' and 'Lng' column for coordinates. They are in decimal degree format, hence we will assume the data is wgs84.

We will next check if there are any NA value in 'Lat' and 'Lng' using is.na function of base R package.

```{r}
any(is.na(hotels$Lat))
```

```{r}
any(is.na(hotels$Lng))
```

We can confirmed that there are no NA values and we can proceed to create a simple feature data frame.

#### 4.2.1.3 Create simple feature data frame from aspatial data

The following code 
1. convert hotels data  into simple feature data frame using st_as_sf feature of sf package.
2. 'Lng' and 'Lat' column are set as the coords arguments.
3. CRS were set as 4326 since we assumed that the data is wgs84
4. st_transform function of sf package then helps to transform the newly created simple feature data frame into svy21
5. tm_shape and tm_dots function of tmap plots the transformed hotels_sf

```{r}
hotels_sf <- st_as_sf(hotels,
                    coords = c("Lng", 
                               "Lat"),
                    crs=4326) %>%
  st_transform(crs = 3414)

tm_shape(hotels_sf)+
  tm_dots()
```

### 4.2.2 Tourism

#### 4.2.2.1 Importing

The following code uses read_csv function of readr package to import Tourism aspatial data sets.

```{r}
tourism <- read_csv("data/aspatial/tourism.csv")
```

#### 4.2.2.2 Examine the file

The following code uses list function of base R to examine the imported file

```{r}
list(tourism)
```

The tourism data contains 107 rows and 17 columns. There are 'LATITUDE' and 'LONGTITUDE' column, as well as 'Lat' and 'Lng' coluumn for coordinates. We will be using 'LATITUDE' and 'LONGTITUDE' column since they are the same. They are in decimal degree format, hence we will assume the data is wgs84.

We will next check if there are any NA value in 'LATITUDE' and 'LONGTITUDE' using is.na function of base R package.

```{r}
any(is.na(tourism$LATITUDE))
```

```{r}
any(is.na(tourism$LATITUDE))
```

There are NA values and we will need to drop them before creating a simple feature data frame.

The following code chunk drop rows of data as long as the LATITUDE is NA.

```{r}
tourism <- tourism %>%
            drop_na("LATITUDE")
```


#### 4.2.2.3 Create simple feature data frame from aspatial data

The following code 
1. convert tourism data into simple feature data frame using st_as_sf feature of sf package.
2. 'LONGTITUDE' and 'LATITUDE' column are set as the coords arguments. 
3. CRS were set as 4326 since we assumed that the data is wgs84
4. st_transform function of sf package then helps to transform the newly created simple feature data frame into svy21
5. tm_shape and tm_dots function of tmap plots the transformed tourism_sf

```{r}

tourism_sf <- st_as_sf(tourism,
                    coords = c("LONGTITUDE", 
                               "LATITUDE"),
                    crs=4326) %>%
  st_transform(crs = 3414)

tm_shape(tourism_sf)+
  tm_dots()
```

### 4.2.3 Airbnb0619

#### 4.2.3.1 Importing

The following code uses read_csv function of readr package to import Airbnb listings30062019.csv aspatial data sets.

```{r}
airbnb0619 <- read_csv("data/aspatial/listings30062019.csv")
```

#### 4.2.3.2 Examine the file

The following code uses list function of base R to examine the imported file

```{r}
list(airbnb0619)
```

The airbnb0619 data contains 8293 rows and 16 columns. There are 'latitude' and 'longitude' column for coordinates. They are in decimal degree format, hence we will assume the data is wgs84.

We will next check if there are any NA value in 'latitude' and 'longitude' using is.na function of base R package.

```{r}
any(is.na(airbnb0619$latitude))
```

```{r}
any(is.na(airbnb0619$longitude))
```

Since there are no NA values, we can proceed to create simple features data frame for airbnb0619

#### 4.2.3.3 Create simple feature data frame from aspatial data

The following code 
1. convert airbnb0619 data into simple feature data frame using st_as_sf feature of sf package.
2. 'longitude' and 'latitude' column are set as the coords arguments. 
3. CRS were set as 4326 since we assumed that the data is wgs84
4. st_transform function of sf package then helps to transform the newly created simple feature data frame into svy21
5. tm_shape and tm_dots function of tmap plots the transformed airbnb0619_sf

```{r}
airbnb0619_sf <- st_as_sf(airbnb0619,
                    coords = c("longitude", 
                               "latitude"),
                    crs=4326) %>%
  st_transform(crs = 3414)

tm_shape(airbnb0619_sf)+
  tm_dots()
```

### 4.2.4 Airbnb0721

#### 4.2.4.1 Importing

The following code uses read_csv function of readr package to import Airbnb listings19072021.csv aspatial data sets.

```{r}
airbnb0721 <- read_csv("data/aspatial/listings19072021.csv")
```

#### 4.2.4.2 Examine the file

The following code uses list function of base R to examine the imported file

```{r}
list(airbnb0721)
```

The airbnb0721 data contains 4252 rows and 16 columns, lesser than airbnb0619. There are 'latitude' and 'longitude' column for coordinates. They are in decimal degree format, hence we will assume the data is wgs84.

We will next check if there are any NA value in 'latitude' and 'longitude' using is.na function of base R package.

```{r}
any(is.na(airbnb0721$latitude))
```

```{r}
any(is.na(airbnb0721$longitude))
```

We can confirm that there are no latitude and longitude with NA value.

#### 4.2.4.3 Create simple feature data frame from aspatial data

The following code 
1. convert airbnb0721 data into simple feature data frame using st_as_sf feature of sf package.
2. 'longitude' and 'latitude' column are set as the coords arguments. 
3. CRS were set as 4326 since we assumed that the data is wgs84
4. st_transform function of sf package then helps to transform the newly created simple feature data frame into svy21
5. tm_shape and tm_dots function of tmap plots the transformed airbnb0721_sf

```{r}
airbnb0721_sf <- st_as_sf(airbnb0721,
                    coords = c("longitude", 
                               "latitude"),
                    crs=4326) %>%
  st_transform(crs = 3414)

tm_shape(airbnb0721_sf)+
  tm_dots()
```

## 4.3 Geospatial Data wrangling

After importing geospatial and aspatial data, we will convert the simple feature data frame to sp’s Spatial* class.

### 4.3.1 Converting sf data frames to sp’s Spatial* class

The following code chunk converts all the geospatial data from simple feature data frame to sp’s Spatial* class using as_Spatial functions of sf package.

```{r}
airbnb0619 <- as_Spatial(airbnb0619_sf)
airbnb0721 <- as_Spatial(airbnb0721_sf)
hotels <- as_Spatial(hotels_sf)
tourism <- as_Spatial(tourism_sf)
mrtlrt <- as_Spatial(mrtlrt_sf)
sg <- as_Spatial(sg_sf)
mpsz <- as_Spatial(mpsz_sf)
```

Check each Spatial* class

```{r}
airbnb0619
```

```{r}
airbnb0721
```

```{r}
hotels
```

```{r}
tourism
```

```{r}
mrtlrt
```

```{r}
sg
```

```{r}
mpsz
```

We can see that all the geospatial data are converted to their respective Spatial* Classes.

### 4.3.2 Converting the Spatial* class into generic sp format

Next, we will convert Spatial* classes into Spatial object so that we can convert it into ppp object.

The following code chunk converts Spatial* classes into generic sp format using as function which sets the classes to their respective sp format. 

```{r}
airbnb0619_sp <- as(airbnb0619, "SpatialPoints")
airbnb0721_sp <- as(airbnb0721, "SpatialPoints")
hotels_sp <- as(hotels, "SpatialPoints")
tourism_sp <- as(tourism, "SpatialPoints")
mrtlrt_sp <- as(mrtlrt, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

### 4.3.3 Converting the generic sp format into spatstat’s ppp format

We will next convert the spatial data into spatstat's ppp format.

The following code chunk converts spatial data into spatstat's ppp format using as function. 

```{r}
airbnb0619_ppp <- as(airbnb0619_sp, "ppp")
airbnb0721_ppp <- as(airbnb0721_sp, "ppp")
hotels_ppp <- as(hotels_sp, "ppp")
tourism_ppp <- as(tourism_sp, "ppp")
mrtlrt_ppp <- as(mrtlrt_sp, "ppp")
```

### 4.3.4 Handling duplicates in ppp object

#### 4.3.4.1 Check for dupliciates duplicates in each ppp object

The following code chunk use any and duplicated function of base R package to check for duplicates, and sum up the count of duplicated points using sum function of base R package and multiplicity function of spatstat package.

```{r}
any(duplicated(airbnb0619_ppp))
sum(multiplicity(airbnb0619_ppp) > 1)
```

From the result above, there are 6 duplicates in airbnb0619_ppp object. 

```{r}
any(duplicated(airbnb0721_ppp))
sum(multiplicity(airbnb0721_ppp) > 1)
```

From the result above, there are 224 duplicates in airbnb0721_ppp object. 

```{r}
any(duplicated(hotels_ppp))
sum(multiplicity(hotels_ppp) > 1)
```

From the result above, there are 10 duplicates in hotels_ppp object.

```{r}
any(duplicated(tourism_ppp))
sum(multiplicity(tourism_ppp) > 1)
```

From the result above, there are 7 duplicates in tourism_ppp object.

```{r}
any(duplicated(mrtlrt_ppp))
sum(multiplicity(mrtlrt_ppp) > 1)
```

From the result above, there are no duplicate in mrtlrt_ppp object.

#### 4.3.4.2 Handling duplicates

We will use jittering to handle duplicates so that we will not lose useful points. Jittering adds a small perturbation to the duplicated points so that they do not occupy the exact same space.

The following code chunk uses rjitter function of spatstat to jitter the ppp object with duplicates.

```{r}
airbnb0619_ppp_jit <- rjitter(airbnb0619_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

airbnb0721_ppp_jit <- rjitter(airbnb0721_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

hotels_ppp_jit <- rjitter(hotels_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

tourism_ppp_jit <- rjitter(tourism_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

#### 4.3.4.3 Check if duplicates still exist

```{r}
any(duplicated(airbnb0619_ppp_jit))
sum(multiplicity(airbnb0619_ppp_jit) > 1)
```

```{r}
any(duplicated(airbnb0721_ppp_jit))
sum(multiplicity(airbnb0721_ppp_jit) > 1)
```

```{r}
any(duplicated(hotels_ppp_jit))
sum(multiplicity(hotels_ppp_jit) > 1)
```

```{r}
any(duplicated(tourism_ppp_jit))
sum(multiplicity(tourism_ppp_jit) > 1)
```

We can confirm that there are no duplicated points after jittering.

### 4.3.5 Create owin objects

We will next create owin object to represent the Singapore boundary.

The following code chunk converts sg SpatialPolygon objects into owin object of spatstat using as function. 

```{r}
sg_owin <- as(sg_sp, "owin")
```

The following code chunk plots sg_owin using plot function of base R package.

```{r}
plot(sg_owin)
```

### 4.3.6 Combining point events object and owin object

The following code chunk combines the point objects of airbnb0619, airbnb0721, hotels, tourism and mrtlrt with the owin object into one ppp object class.

```{r}
airbnb0619SG_ppp = airbnb0619_ppp_jit[sg_owin]
airbnb0721SG_ppp = airbnb0721_ppp_jit[sg_owin]
hotelsSG_ppp = hotels_ppp_jit[sg_owin]
tourismSG_ppp = tourism_ppp_jit[sg_owin]
mrtlrtSG_ppp = mrtlrt_ppp[sg_owin]
```

The following code chunk plots the ppp object class using plot function of base R package.

```{r}
plot(airbnb0619SG_ppp)
```

```{r}
plot(airbnb0721SG_ppp)
```

```{r}
plot(hotelsSG_ppp)
```

```{r}
plot(tourismSG_ppp)
```

```{r}
plot(mrtlrtSG_ppp)
```


# 5 Section A: Airbnb Distribution in 2019

## 5.1 Exploratory Spatial Data Analysis

### 5.1.1 Kernel density maps

#### 5.1.1.1 Airbnb0619 listings

##### 5.1.1.1.1 Compute kernel density estimation with automatic bandwidth selection method

1. The following code chunk computes kernel density using density function of spatstat package.
2. The code first takes in the ppp object variable.
3. bw.diggle automatic bandwidth method to detect a tight cluster
4. smoothing kernel is set to default gaussian method 

```{r}
kde_airbnb0619SG_bw <- density(airbnb0619SG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")
```

The following code chunk plots the kde_airbnb0619SG_bw using plot function of base R

```{r}
plot(kde_airbnb0619SG_bw)
```

The density values of the output range is  too small because the default unit of measurement of svy21 is in meter. As a result, the density values computed is in “number of points per square meter”.

##### 5.1.1.1.2 Rescalling KDE values

The following code chunk rescale the KDE value of airbnb0619SG_ppp using rescale function of spatstat from meter to kilometer.

```{r}
airbnb0619SG_ppp.km <- rescale(airbnb0619SG_ppp, 1000, "km")
```

##### 5.1.1.1.3 Different automatic bandwidth methods

The following code chunk calculate the bandwidth using different methods

```{r}
bw.CvL(airbnb0619SG_ppp.km)
```

```{r}
bw.scott(airbnb0619SG_ppp.km)
```

```{r}
bw.ppl(airbnb0619SG_ppp.km)
```

```{r}
bw.diggle(airbnb0619SG_ppp.km)
```

From the result above, since bw.diggle has the lowest sigma, we will be using it for the other ppp object in order to detect close cluster.

##### 5.1.1.1.4 KDE Map and Point Map

1. The following code chunk computes kernel density using density function of spatstat package.
2. The code first takes in the ppp object variable in km.
3. bw.diggle automatic bandwidth method to detect a tight cluster
4. smoothing kernel is set to default gaussian method
5. plot function of base R package and par function of graphics package is used to plot kde map and ppp point map side by side

```{r}
kde_airbnb0619SG_bw.km <- density(airbnb0619SG_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")


par(mfrow=c(1,2))
plot(kde_airbnb0619SG_bw.km)
plot(airbnb0619SG_ppp)
```


##### 5.1.1.1.5 Convert KDE output into grid object

The following code chunk converts KDE output into grid object using as.SpatialGridDataFrame.im function of maptools, and plot the grid object using spplot function of raster package.

```{r}
gridded_kde_airbnb0619SG_bw <- as.SpatialGridDataFrame.im(kde_airbnb0619SG_bw.km)
spplot(gridded_kde_airbnb0619SG_bw)
```

##### 5.1.1.1.6 Convert grid object into raster

The following code chunk converts grid object into raster using raster function of raster package

```{r}
kde_airbnb0619SG_bw_raster <- raster(gridded_kde_airbnb0619SG_bw)
kde_airbnb0619SG_bw_raster
```

From the result above, the crs is NA

##### 5.1.1.1.7 Assign project system

The following code chunk assign CRS to kde_airbnb0619SG_bw_raster

```{r}
projection(kde_airbnb0619SG_bw_raster) <- CRS("+init=EPSG:3414")
kde_airbnb0619SG_bw_raster
```

#### 5.1.1.2 Hotels

##### 5.1.1.2.1 Rescale KDE values

The following code chunk rescale the KDE value of hotelsSG_ppp using rescale function of spatstat from meter to kilometer.

```{r}
hotelsSG_ppp.km <- rescale(hotelsSG_ppp, 1000, "km")
```

##### 5.1.1.2.2 KDE Map and Point Map

1. The following code chunk computes kernel density using density function of spatstat package.
2. The code first takes in the ppp object variable in km.
3. bw.diggle automatic bandwidth method to detect a tight cluster
4. smoothing kernel is set to default gaussian method 
5. plot function of base R package and par function of graphics package is used to plot kde map and ppp point map side by side

```{r}
kde_hotelsSG_bw.km <- density(hotelsSG_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

par(mfrow=c(1,2))
plot(kde_hotelsSG_bw.km, main = "KDE Map")
plot(hotelsSG_ppp, main="Point Map")
```

##### 5.1.1.2.3 Convert KDE output into grid object

The following code chunk converts KDE output into grid object using as.SpatialGridDataFrame.im function of maptools, and plot the grid object using spplot function of raster package.

```{r}
gridded_kde_hotelsSG_bw <- as.SpatialGridDataFrame.im(kde_hotelsSG_bw.km)
spplot(gridded_kde_hotelsSG_bw)
```

##### 5.1.1.2.4 Convert grid object into raster and assign project system

The following code chunk converts grid object into raster using raster function of raster package

```{r}
kde_hotelsSG_bw_raster <- raster(gridded_kde_hotelsSG_bw)
projection(kde_hotelsSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_hotelsSG_bw_raster
```

#### 5.1.1.3 MRTLRT

##### 5.1.1.3.1 Rescale KDE values

The following code chunk rescale the KDE value of mrtlrtSG_ppp using rescale function of spatstat from meter to kilometer.

```{r}
mrtlrtSG_ppp.km <- rescale(mrtlrtSG_ppp, 1000, "km")
```

##### 5.1.1.3.2 KDE Map and Point Map

1. The following code chunk computes kernel density using density function of spatstat package.
2. The code first takes in the ppp object variable in km.
3. bw.diggle automatic bandwidth method to detect a tight cluster
4. smoothing kernel is set to default gaussian method 
5. plot function of base R package and par function of graphics package is used to plot kde map and ppp point map side by side

```{r}
kde_mrtlrtSG_bw.km <- density(mrtlrtSG_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

par(mfrow=c(1,2))
plot(kde_mrtlrtSG_bw.km, main = "KDE Map")
plot(mrtlrtSG_ppp, main="Point Map")
```

##### 5.1.1.3.3 Convert KDE output into grid object

The following code chunk converts KDE output into grid object using as.SpatialGridDataFrame.im function of maptools, and plot the grid object using spplot function of raster package.

```{r}
gridded_kde_mrtlrtSG_bw <- as.SpatialGridDataFrame.im(kde_mrtlrtSG_bw.km)
spplot(gridded_kde_mrtlrtSG_bw)
```

##### 5.1.1.3.4 Convert grid object into raster and assign project system

The following code chunk converts grid object into raster using raster function of raster package

```{r}
kde_mrtlrtSG_bw_raster <- raster(gridded_kde_mrtlrtSG_bw)
projection(kde_mrtlrtSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_mrtlrtSG_bw_raster
```

#### 5.1.1.4 Tourist Attractions

##### 5.1.1.4.1 Rescale KDE values

The following code chunk rescale the KDE value of mrtlrtSG_ppp using rescale function of spatstat from meter to kilometer.

```{r}
tourismSG_ppp.km <- rescale(tourismSG_ppp, 1000, "km")
```

##### 5.1.1.4.2 KDE Map and Point Map

1. The following code chunk computes kernel density using density function of spatstat package.
2. The code first takes in the ppp object variable in km.
3. bw.diggle automatic bandwidth method to detect a tight cluster
4. smoothing kernel is set to default gaussian method 
5. plot function of base R package and par function of graphics package is used to plot kde map and ppp point map side by side

```{r}
kde_tourismSG_bw.km <- density(tourismSG_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

par(mfrow=c(1,2))
plot(kde_tourismSG_bw.km, main = "KDE Map")
plot(tourismSG_ppp, main="Point Map")
```

##### 5.1.1.4.3 Convert KDE output into grid object

The following code chunk converts KDE output into grid object using as.SpatialGridDataFrame.im function of maptools, and plot the grid object using spplot function of raster package.

```{r}
gridded_kde_tourismSG_bw <- as.SpatialGridDataFrame.im(kde_tourismSG_bw.km)
spplot(gridded_kde_tourismSG_bw)
```

##### 5.1.1.4.4 Convert grid object into raster and assign project system

The following code chunk converts grid object into raster using raster function of raster package

```{r}
kde_tourismSG_bw_raster <- raster(gridded_kde_tourismSG_bw)
projection(kde_tourismSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_tourismSG_bw_raster
```


#### 5.1.1.5 Comparison between Airbnb0619, Hotels, MRTLRT and Tourist Attractions

##### 5.1.1.5.1 KDE Map Comparison

The following code chunk plots all the KDE map of Airbnb0619, Hotels, MRTLRT and Tourist Attractions side by side using par functions of graphics package

```{r}
par(mfrow=c(2,2))
plot(kde_airbnb0619SG_bw.km, main = "Airbnb 2019")
plot(kde_hotelsSG_bw.km, main = "Hotel")
plot(kde_mrtlrtSG_bw.km, main = "Mrt")
plot(kde_tourismSG_bw.km, main = "Tourist Attraction")
```

From the above, we can see that the higher density (in red and yellow) for airbnb, hotel, mrt and tourish attraction are concentrated around the south east area below central region.

##### 5.1.1.5.2 Central Region

We will look into the central region

###### 5.1.1.5.2.1 Exploration with OpenStreetMap

The following code chunk first extract airbnb in the central region from airbnb0619_sf region.

tmap_mode is set to "view" for interaction mode and tm_basemap is set to "OpenStreetMap"

tm_shape of centralairbnb0619, hotels_sf, mrtlrt_sf and tourism_sf were plotted using tm_shape and tm_dots of tmap packages.

colour of dots using tm_dots and the following are the color representation:

- black: Airbnb listings in central region

- green: hotels

- blue: mrtlrt

- red: tourist attractions.


```{r}
centralairbnb0619 <- airbnb0619_sf[airbnb0619_sf$neighbourhood_group == "Central Region",]

tmap_mode("view")
tm_basemap("OpenStreetMap") +
tm_shape(centralairbnb0619) + 
  tm_dots(col="black", size = 0.02)+
tm_shape(hotels_sf) + 
  tm_dots(col="green", size = 0.02)+
tm_shape(mrtlrt_sf) + 
  tm_dots(col="blue", size = 0.02)+
tm_shape(tourism_sf) + 
  tm_dots(col="red", size = 0.02)
```

From the map above, we can see that there are many Airbnbs located in Kallang, River Valley, Novena an Rochor.

```{r}
tmap_mode("plot")
```

###### 5.1.1.5.2.2 Comparing Spatial Point Patterns using KDE

Extract Study Area

The following code chunk extracts the target areas.

```{r}
kl = mpsz[mpsz@data$PLN_AREA_N == "KALLANG",]
rv = mpsz[mpsz@data$PLN_AREA_N == "RIVER VALLEY",]
nv = mpsz[mpsz@data$PLN_AREA_N == "NOVENA",]
rc = mpsz[mpsz@data$PLN_AREA_N == "ROCHOR",]
```

Plot target areas

The following code chunk plots the target areas side by side in 2x2 using par function of graphics package and plot functions of base R package. 

```{r}
par(mfrow=c(2,2))
plot(kl, main = "Kallang")
plot(rv, main = "River Valley")
plot(nv, main = "Novena")
plot(rc, main = "Rochor")
```

Convert Spatial Polygons Data Fraeme into generic SP format

The following code chunk converts the four SpatialPolygonsDataFrame layers into generic spatialpolygons layers.

```{r}
kl_sp = as(kl, "SpatialPolygons")
rv_sp = as(rv, "SpatialPolygons")
nv_sp = as(nv, "SpatialPolygons")
rc_sp = as(rc, "SpatialPolygons")
```

Create owin object for each target area

The following code chunk converst the SpatialPolygons objects into owin objects required by spatstat.

```{r}
kl_owin = as(kl_sp, "owin")
rv_owin = as(rv_sp, "owin")
nv_owin = as(nv_sp, "owin")
rc_owin = as(rc_sp, "owin")
```

Combine airnbnb0619, hotels, mrt, tourist attractions points with owin of each target area

The following code chunk extracts airnbnb0619, hotels, mrt and tourist attractions within each specific region.

```{r}
airbnb0619_kl_ppp = airbnb0619_ppp_jit[kl_owin]
airbnb0619_rv_ppp = airbnb0619_ppp_jit[rv_owin]
airbnb0619_nv_ppp = airbnb0619_ppp_jit[nv_owin]
airbnb0619_rc_ppp = airbnb0619_ppp_jit[rc_owin]

hotels_kl_ppp = hotels_ppp_jit[kl_owin]
hotels_rv_ppp = hotels_ppp_jit[rv_owin]
hotels_nv_ppp = hotels_ppp_jit[nv_owin]
hotels_rc_ppp = hotels_ppp_jit[rc_owin]

mrtlrt_kl_ppp = mrtlrt_ppp[kl_owin]
mrtlrt_rv_ppp = mrtlrt_ppp[rv_owin]
mrtlrt_nv_ppp = mrtlrt_ppp[nv_owin]
mrtlrt_rc_ppp = mrtlrt_ppp[rc_owin]

tourism_kl_ppp = tourism_ppp_jit[kl_owin]
tourism_rv_ppp = tourism_ppp_jit[rv_owin]
tourism_nv_ppp = tourism_ppp_jit[nv_owin]
tourism_rc_ppp = tourism_ppp_jit[rc_owin]
```

Rescale from m to km

The following code chunk transform the unit from m to km using rescale function.

```{r}
airbnb0619_kl_ppp.km = rescale(airbnb0619_kl_ppp, 1000, "km")
airbnb0619_rv_ppp.km = rescale(airbnb0619_rv_ppp, 1000, "km")
airbnb0619_nv_ppp.km = rescale(airbnb0619_nv_ppp, 1000, "km")
airbnb0619_rc_ppp.km = rescale(airbnb0619_rc_ppp, 1000, "km")

hotels_kl_ppp.km = rescale(hotels_kl_ppp, 1000, "km")
hotels_rv_ppp.km = rescale(hotels_rv_ppp, 1000, "km")
hotels_nv_ppp.km = rescale(hotels_nv_ppp, 1000, "km")
hotels_rc_ppp.km = rescale(hotels_rc_ppp, 1000, "km")

mrtlrt_kl_ppp.km = rescale(mrtlrt_kl_ppp, 1000, "km")
mrtlrt_rv_ppp.km = rescale(mrtlrt_rv_ppp, 1000, "km")
mrtlrt_nv_ppp.km = rescale(mrtlrt_nv_ppp, 1000, "km")
mrtlrt_rc_ppp.km = rescale(mrtlrt_rc_ppp, 1000, "km")

tourism_kl_ppp.km = rescale(tourism_kl_ppp, 1000, "km")
tourism_rv_ppp.km = rescale(tourism_rv_ppp, 1000, "km")
tourism_nv_ppp.km = rescale(tourism_nv_ppp, 1000, "km")
tourism_rc_ppp.km = rescale(tourism_rc_ppp, 1000, "km")
```

Airbnb comparison by neighbourhood

The following code chunk plots airbnb within each neigbourhood using plot and density functions with sigma set to bw.diggle and kernel set to gaussian.

```{r}
par(mfrow=c(2,2))
plot(density(airbnb0619_kl_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Kallang")
plot(density(airbnb0619_rv_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="River Valley")
plot(density(airbnb0619_nv_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Novena")
plot(density(airbnb0619_rc_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Rochor")
```

Analysis: 

Comparison within Kallang

The following code chunk plots airbnb, hotel, mrt and tourist attraction in Kallang.

```{r}
par(mfrow=c(2,2))
plot(airbnb0619_kl_ppp.km, main = "Airbnb")
plot(hotels_kl_ppp.km, main = "Hotel")
plot(mrtlrt_kl_ppp.km, main = "Mrt")
plot(tourism_kl_ppp.km, main = "Tourist Attraction")
```

Analysis: Even though there are no tourist attraction and hotels, there are still many airbnb in Kallang. There are a few MRT stations in Kallang, which may be the reason to the large number of airbnb.

Comparison within River Valley

The following code chunk plots airbnb, hotel, mrt and tourist attraction in River Valley.

```{r}
par(mfrow=c(2,2))
plot(airbnb0619_rv_ppp.km, main = "Airbnb")
plot(hotels_rv_ppp.km, main = "Hotel")
plot(mrtlrt_rv_ppp.km, main = "Mrt")
plot(tourism_rv_ppp.km, main = "Tourist Attraction")
```

Analysis: Even though there are no tourist attraction and Mrt in River Valley, there are still many airbnb. There are only one hotel and it may be the reason for the large number of airbnb in the area.

Comparison within Novena

The following code chunk plots airbnb, hotel, mrt and tourist attraction in Novena.

```{r}
par(mfrow=c(2,2))
plot(airbnb0619_nv_ppp.km, main = "Airbnb")
plot(hotels_nv_ppp.km, main = "Hotel")
plot(mrtlrt_nv_ppp.km, main = "Mrt")
plot(tourism_nv_ppp.km, main = "Tourist Attraction")
```
Analysis: Even though there are a few hotel in Novena, the number of airbnb are large. The accessibility of the two MRT and one tourist attraction can be the reason to the large number of airbnb in Novena.

Comparison within Rochor

The following code chunk plots airbnb, hotel, mrt and tourist attraction in Rochor.

```{r}
par(mfrow=c(2,2))
plot(airbnb0619_rc_ppp.km, main = "Airbnb")
plot(hotels_rc_ppp.km, main = "Hotel")
plot(mrtlrt_rc_ppp.km, main = "Mrt")
plot(tourism_rc_ppp.km, main = "Tourist Attraction")
```

Analysis: In Rochor, there are still many airbnb even though many hotels are already available. The accessilibty to a few MRTs and tourist attractions may be the reason that leads to a large number of airbnb in Rochor.

Overall, we can conclude that there are no direct relationship between airbnb and hotel, mrt and tourist attraction.

## 5.2 Second-order Spatial Point Patterns Analysis

### 5.2.1 Nearest Neighbour Analysis

#### 5.2.1.1 Hypothesis

we will perform the Clark-Evans test of aggregation for a spatial point pattern by using clarkevans.test() of statspat.

Test Hypothesis:

H0 = The distribution of airbnb are randomly distributed.

H1 = The distribution of airbnb are not randomly distributed.

The 95% confident interval will be used.

#### 5.2.1.2 Testing and Analysis

##### 5.2.1.2.1 Testing spatial point patterns using Clark and Evans Test

```{r}
clarkevans.test(airbnb0619SG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

  Clark-Evans test
	No edge correction
	Monte Carlo test based on 99 simulations of CSR with fixed n

data:  airbnb0619SG_ppp
R = 0.36169, p-value = 0.01
alternative hypothesis: clustered (R < 1)

##### 5.2.1.2.2 Clark and Evans Test: Kallang planning area

```{r}
clarkevans.test(airbnb0619_kl_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: 

##### 5.2.1.2.3 Clark and Evans Test: River Valley planning area

```{r}
clarkevans.test(airbnb0619_rv_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: 

##### 5.2.1.2.4 Clark and Evans Test: Novena planning area

```{r}
clarkevans.test(airbnb0619_nv_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: 

##### 5.2.1.2.5 Clark and Evans Test: Rochor planning area

```{r}
clarkevans.test(airbnb0619_rc_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: 


### G-FUNCTION

#### Hypothesis

#### Testing and Analysis

### F-FUNCTION

#### Hypothesis

#### Testing and Analysis

### K-FUNCTION

#### Hypothesis

#### Testing and Analysis

### L-FUNCTION

#### Hypothesis

#### Testing and Analysis


# 6 Section B: Impact of COVID-19

## 6.1 Exploratory Spatial Data Analysis

### 6.1.1 Kernel density maps

#### 6.1.1.1 Airbnb0721 listings

##### 6.1.1.2.1 Rescale KDE values

The following code chunk rescale the KDE value of airbnb0721SG_ppp using rescale function of spatstat from meter to kilometer.

```{r}
airbnb0721SG_ppp.km <- rescale(airbnb0721SG_ppp, 1000, "km")
```

##### 6.1.1.2.2 KDE Map

1. The following code chunk computes kernel density using density function of spatstat package.
2. The code first takes in the ppp object variable in km.
3. bw.diggle automatic bandwidth method to detect a tight cluster
4. smoothing kernel is set to default gaussian method
5. plot function of base R package to plot the kde map

```{r}
kde_airbnb0721SG_bw.km <- density(airbnb0721SG_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

plot(kde_airbnb0721SG_bw.km, main = "KDE Map")
```

##### 6.1.1.2.3 Convert KDE output into grid object.

The following code chunk converts KDE output into grid object using as.SpatialGridDataFrame.im function of maptools, and plot the grid object using spplot function of raster package.

```{r}
gridded_kde_airbnb0721SG_bw <- as.SpatialGridDataFrame.im(kde_airbnb0721SG_bw.km)
spplot(gridded_kde_airbnb0721SG_bw)
```

##### 6.1.1.2.4 Convert grid object into raster and assign project system

The following code chunk converts grid object into raster using raster function of raster package

```{r}
kde_airbnb0721SG_bw_raster <- raster(gridded_kde_airbnb0721SG_bw)
projection(kde_airbnb0721SG_bw_raster) <- CRS("+init=EPSG:3414")
kde_airbnb0721SG_bw_raster
```


#### 6.1.1.2 Comparison between Airbnb0619 and Airbnb0721

##### 6.1.1.2.1 KDE Map Comparison

```{r}
par(mfrow=c(1,2))
plot(kde_airbnb0619SG_bw.km, main = "Airbnb 2019")
plot(kde_airbnb0721SG_bw.km, main = "Airbnb 2021")
```

From the KDE map, the south-east region still have higher density in 2021. However, we notice the density of Airbnb have reduced as seen from the scale. It reduced from range 0-1500 in 2019 to range 0-1000 in 2021. We can conclude that Covid-19 did impact the Airbnb listing in Singapore from 2019 to 2021.

#### 6.1.1.3 Comparison between Airbnb0619 and Airbnb0721 by Room Type

##### 6.1.1.3.1 Data Wrangling

###### 6.1.1.3.1.1 Determine the unique room type

The following code checks the unique room type using unique features of base R

```{r}
unique(airbnb0619_sf$room_type)
```

We have determined that the unique room types are Private room, Entire home/apt and Shared room for Airbnb 2021.


```{r}
unique(airbnb0721_sf$room_type)
```

We have determined that the unique room types are Private room, Entire home/apt, Shared room and Hotel room for Airbnb 2021.

###### 6.1.1.3.1.2 Filter each room type from SpatialPointsData Frama

```{r}
airbnb0619_pte = airbnb0619[airbnb0619@data$room_type == "Private room",]
airbnb0619_ent = airbnb0619[airbnb0619@data$room_type == "Entire home/apt",]
airbnb0619_sh = airbnb0619[airbnb0619@data$room_type == "Shared room",]
airbnb0721_pte = airbnb0721[airbnb0721@data$room_type == "Private room",]
airbnb0721_ent = airbnb0721[airbnb0721@data$room_type == "Entire home/apt",]
airbnb0721_sh = airbnb0721[airbnb0721@data$room_type == "Shared room",]
airbnb0721_ht = airbnb0721[airbnb0721@data$room_type == "Hotel room",]
```


```{r}
airbnb0721_sh
```

Notice that the data are in SpatialPointsData Frame format

###### 6.1.1.3.1.3 Converting the Spatial* class into generic sp format

Next, we will convert Spatial* classes into Spatial object so that we can convert it into ppp object.

The following code chunk converts Spatial* classes into generic sp format using as function which sets the classes to their respective sp format. 

```{r}
airbnb0619_pte_sp <- as(airbnb0619_pte, "SpatialPoints")
airbnb0619_ent_sp <- as(airbnb0619_ent, "SpatialPoints")
airbnb0619_sh_sp <- as(airbnb0619_sh, "SpatialPoints")
airbnb0721_pte_sp <- as(airbnb0721_pte, "SpatialPoints")
airbnb0721_ent_sp <- as(airbnb0721_ent, "SpatialPoints")
airbnb0721_sh_sp <- as(airbnb0721_sh, "SpatialPoints")
airbnb0721_ht_sp <- as(airbnb0721_ht, "SpatialPoints")
```


###### 6.1.1.3.1.4 Converting the generic sp format into spatstat’s ppp format

We will next convert the spatial data into spatstat's ppp format.

The following code chunk converts spatial data into spatstat's ppp format using as function. 

```{r}
airbnb0619_pte_ppp <- as(airbnb0619_pte_sp, "ppp")
airbnb0619_ent_ppp <- as(airbnb0619_ent_sp, "ppp")
airbnb0619_sh_ppp <- as(airbnb0619_sh_sp, "ppp")
airbnb0721_pte_ppp <- as(airbnb0721_pte_sp, "ppp")
airbnb0721_ent_ppp <- as(airbnb0721_ent_sp, "ppp")
airbnb0721_sh_ppp <- as(airbnb0721_sh_sp, "ppp")
airbnb0721_ht_ppp <- as(airbnb0721_ht_sp, "ppp")
```

###### 6.1.1.3.1.5 Check for duplicates in each ppp's objects

The following code chunk use any and duplicated function of base R package to check for duplicates, and sum up the count of duplicated points using sum function of base R package and multiplicity function of spatstat package.

```{r}
any(duplicated(airbnb0619_pte_ppp))
sum(multiplicity(airbnb0619_pte_ppp) > 1)
```

```{r}
any(duplicated(airbnb0619_ent_ppp))
sum(multiplicity(airbnb0619_ent_ppp) > 1)
```

```{r}
any(duplicated(airbnb0619_sh_ppp))
sum(multiplicity(airbnb0619_sh_ppp) > 1)
```

```{r}
any(duplicated(airbnb0721_pte_ppp))
sum(multiplicity(airbnb0721_pte_ppp) > 1)
```

```{r}
any(duplicated(airbnb0721_ent_ppp))
sum(multiplicity(airbnb0721_ent_ppp) > 1)
```

```{r}
any(duplicated(airbnb0721_sh_ppp))
sum(multiplicity(airbnb0721_sh_ppp) > 1)
```

```{r}
any(duplicated(airbnb0721_ht_ppp))
sum(multiplicity(airbnb0721_ht_ppp) > 1)
```

###### 6.1.1.3.1.6 Handling Duplicates

We will use jittering to handle duplicates so that we will not lose useful points. Jittering adds a small perturbation to the duplicated points so that they do not occupy the exact same space.

The following code chunk uses rjitter function of spatstat to jitter the ppp object with duplicates.

```{r}
airbnb0619_ent_ppp_jit <- rjitter(airbnb0619_ent_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

airbnb0619_sh_ppp_jit <- rjitter(airbnb0619_sh_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

airbnb0721_pte_ppp_jit <- rjitter(airbnb0721_pte_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

airbnb0721_ent_ppp_jit <- rjitter(airbnb0721_ent_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

airbnb0721_sh_ppp_jit <- rjitter(airbnb0721_sh_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

airbnb0721_ht_ppp_jit <- rjitter(airbnb0721_ht_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

###### 6.1.1.3.1.7 Check if duplicates still exist

```{r}
any(duplicated(airbnb0619_ent_ppp_jit))
sum(multiplicity(airbnb0619_ent_ppp_jit) > 1)
```

```{r}
any(duplicated(airbnb0619_sh_ppp_jit))
sum(multiplicity(airbnb0619_sh_ppp_jit) > 1)
```

```{r}
any(duplicated(airbnb0721_pte_ppp_jit))
sum(multiplicity(airbnb0721_pte_ppp_jit) > 1)
```

```{r}
any(duplicated(airbnb0721_ent_ppp_jit))
sum(multiplicity(airbnb0721_ent_ppp_jit) > 1)
```

```{r}
any(duplicated(airbnb0721_sh_ppp_jit))
sum(multiplicity(airbnb0721_sh_ppp_jit) > 1)
```

```{r}
any(duplicated(airbnb0721_ht_ppp_jit))
sum(multiplicity(airbnb0721_ht_ppp_jit) > 1)
```
We can confirm that there are no duplicated points after jittering.



###### 6.1.1.3.1.8 Combining point events object and owin object

The following code chunk combines the point objects with the owin object into one ppp object class.

```{r}
airbnb0619SG_pte_ppp = airbnb0619_pte_ppp[sg_owin]
airbnb0619SG_ent_ppp = airbnb0619_ent_ppp_jit[sg_owin]
airbnb0619SG_sh_ppp = airbnb0619_sh_ppp_jit[sg_owin]
airbnb0721SG_pte_ppp = airbnb0721_pte_ppp_jit[sg_owin]
airbnb0721SG_ent_ppp = airbnb0721_ent_ppp_jit[sg_owin]
airbnb0721SG_sh_ppp = airbnb0721_sh_ppp_jit[sg_owin]
airbnb0721SG_ht_ppp = airbnb0721_ht_ppp_jit[sg_owin]
```

###### 6.1.1.3.1.9 Rescale KDE values

The following code chunk rescale the KDE value of mrtlrtSG_ppp using rescale function of spatstat from meter to kilometer.

```{r}
airbnb0619SG_pte_ppp.km <- rescale(airbnb0619SG_pte_ppp, 1000, "km")
airbnb0619SG_ent_ppp.km <- rescale(airbnb0619SG_ent_ppp, 1000, "km")
airbnb0619SG_sh_ppp.km <- rescale(airbnb0619SG_sh_ppp, 1000, "km")
airbnb0721SG_pte_ppp.km <- rescale(airbnb0721SG_pte_ppp, 1000, "km")
airbnb0721SG_ent_ppp.km <- rescale(airbnb0721SG_ent_ppp, 1000, "km")
airbnb0721SG_sh_ppp.km <- rescale(airbnb0721SG_sh_ppp, 1000, "km")
airbnb0721SG_ht_ppp.km <- rescale(airbnb0721SG_ht_ppp, 1000, "km")
```

###### 6.1.1.3.1.9 KDE

1. The following code chunk computes kernel density using density function of spatstat package.
2. The code first takes in the ppp object variable in km.
3. bw.diggle automatic bandwidth method to detect a tight cluster
4. smoothing kernel is set to default gaussian method

```{r}
kde_airbnb0619SG_pte_bw.km <- density(airbnb0619SG_pte_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

kde_airbnb0619SG_ent_bw.km <- density(airbnb0619SG_ent_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

kde_airbnb0619SG_sh_bw.km <- density(airbnb0619SG_sh_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

kde_airbnb0721SG_pte_bw.km <- density(airbnb0721SG_pte_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

kde_airbnb0721SG_ent_bw.km <- density(airbnb0721SG_ent_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

kde_airbnb0721SG_sh_bw.km <- density(airbnb0721SG_sh_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")

kde_airbnb0721SG_ht_bw.km <- density(airbnb0721SG_ht_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")
```

###### 6.1.1.3.1.10 Convert KDE output into grid object

The following code chunk converts KDE output into grid object using as.SpatialGridDataFrame.im function of maptools

```{r}
gridded_kde_airbnb0619SG_pte_bw <- as.SpatialGridDataFrame.im(kde_airbnb0619SG_pte_bw.km)
gridded_kde_airbnb0619SG_ent_bw <- as.SpatialGridDataFrame.im(kde_airbnb0619SG_ent_bw.km)
gridded_kde_airbnb0619SG_sh_bw <- as.SpatialGridDataFrame.im(kde_airbnb0619SG_sh_bw.km)
gridded_kde_airbnb0721SG_pte_bw <- as.SpatialGridDataFrame.im(kde_airbnb0721SG_pte_bw.km)
gridded_kde_airbnb0721SG_ent_bw <- as.SpatialGridDataFrame.im(kde_airbnb0721SG_ent_bw.km)
gridded_kde_airbnb0721SG_sh_bw <- as.SpatialGridDataFrame.im(kde_airbnb0721SG_sh_bw.km)
gridded_kde_airbnb0721SG_ht_bw <- as.SpatialGridDataFrame.im(kde_airbnb0721SG_ht_bw.km)
```

###### 6.1.1.3.1.11 Convert grid object into raster and assign project system

The following code chunk converts grid object into raster using raster function of raster package

```{r}
kde_airbnb0619SG_pte_bw_raster <- raster(gridded_kde_airbnb0619SG_pte_bw)
projection(kde_airbnb0619SG_pte_bw_raster) <- CRS("+init=EPSG:3414")

kde_airbnb0619SG_ent_bw_raster <- raster(gridded_kde_airbnb0619SG_ent_bw)
projection(kde_airbnb0619SG_ent_bw_raster) <- CRS("+init=EPSG:3414")

kde_airbnb0619SG_sh_bw_raster <- raster(gridded_kde_airbnb0619SG_sh_bw)
projection(kde_airbnb0619SG_sh_bw_raster) <- CRS("+init=EPSG:3414")

kde_airbnb0721SG_pte_bw_raster <- raster(gridded_kde_airbnb0721SG_pte_bw)
projection(kde_airbnb0721SG_pte_bw_raster) <- CRS("+init=EPSG:3414")

kde_airbnb0721SG_ent_bw_raster <- raster(gridded_kde_airbnb0721SG_ent_bw)
projection(kde_airbnb0721SG_ent_bw_raster) <- CRS("+init=EPSG:3414")

kde_airbnb0721SG_sh_bw_raster <- raster(gridded_kde_airbnb0721SG_sh_bw)
projection(kde_airbnb0721SG_sh_bw_raster) <- CRS("+init=EPSG:3414")

kde_airbnb0721SG_ht_bw_raster <- raster(gridded_kde_airbnb0721SG_ht_bw)
projection(kde_airbnb0721SG_ht_bw_raster) <- CRS("+init=EPSG:3414")
```



##### 6.1.1.3.2 KDE Map Comparison by Room type and Year

The following code chunks plot KDE map of Airbnb 2019 and 2021 by each room types side by side.

###### 6.1.1.3.2.1 Private Room

```{r}
par(mfrow=c(1,2))
plot(kde_airbnb0619SG_pte_bw.km, main = "Airbnb 2019 Private Room")
plot(kde_airbnb0721SG_pte_bw.km, main = "Airbnb 2021 Private Room")
```

Analysis: There is a slight decrease in private room airbnb from 2019 to 2021 as seen in 2021 that there are lesser red point. The scale of 2019 to 2021 descreased slightly.

###### 6.1.1.3.2.2 Entire Home/Apt

```{r}
par(mfrow=c(1,2))
plot(kde_airbnb0619SG_ent_bw.km, main = "Airbnb 2019 Entire Home/Apt")
plot(kde_airbnb0721SG_ent_bw.km, main = "Airbnb 2021 Entire Home/Apt")
```

Analysis: The map shows high density of airbnb in south-east region near central for both 2019 and 2021 as seen from the red marks on both map. However, there is a huge decrease in the number of Entire Home/Apt airbnb from 2019 to 2021 as seen on the scale. It decreased from 0-1000 in 2019 to 0-250 in 2021.

###### 6.1.1.3.2.3 Shared Room

```{r}
par(mfrow=c(1,2))
plot(kde_airbnb0619SG_sh_bw.km, main = "Airbnb 2019 Shared Room")
plot(kde_airbnb0721SG_sh_bw.km, main = "Airbnb 2021 Shared Room")
```

Analysis: The map shows high density of airbnb in south-east region near central for both 2019 and 2021 as seen from the red marks on both map. However, there is a significant decrease in the number of Shared Room from 2019 to 2021 as seen on the scale. It ranges from 0-400 in 2019 and decrease to 0-80 in 2021.

###### 6.1.1.3.2.4 Hotel Room

```{r}
plot(kde_airbnb0721SG_ht_bw.km, main = "Airbnb 2021 Hotel Room")
```

Analyis: Hotel room is a new room type in 2021.

Overall, we can conclude that Covid-19 are likely to have impact on the type of Airbnb room listed in Singapore from 2019 to 2021. It is seen from the significant decrease in Entire Home/Apt and Shared Room type, and this is likely because the owner do not wish to rent out the house anymore, or that there are lesser to no tourist. In addition, there is a new room type which is the Hotel Room, and this because hotel are required for quarantine purpose.

##### 6.1.1.3.3 Exploration with Openstreetmap

```{r}
tmap_mode("view")

airbnb0619map <- tm_basemap("OpenStreetMap")+
                  tm_shape(airbnb0619_sf) +
                  tm_dots(col = 'room_type', palette=c(entire='blue', private='red', shared='green'), size = 0.02)

airbnb0721map <- tm_basemap("OpenStreetMap")+
                  tm_shape(airbnb0721_sf) +
                  tm_dots(col = 'room_type', palette=c(entire='blue', hotel='yellow', private='red', shared='green'), size = 0.02)

tmap_arrange(airbnb0619map, airbnb0721map)
```

```{r}
tmap_mode("plot")
```

## 6.2 Second-order Spatial Point Patterns Analysis

### 6.2.1 Nearest Neighbour Analysis (CONTINUE FROM HERE)

#### 6.2.1.1 Hypothesis

we will perform the Clark-Evans test of aggregation for a spatial point pattern by using clarkevans.test() of statspat.

Test Hypothesis:

H0 = The distribution of airbnb are randomly distributed.

H1 = The distribution of airbnb are not randomly distributed.

The 95% confident interval will be used.

#### 6.2.1.2 Testing and Analysis

##### 6.2.1.2.1 Testing spatial point patterns using Clark and Evans Test

```{r}
clarkevans.test(airbnb0619SG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

  Clark-Evans test
	No edge correction
	Monte Carlo test based on 99 simulations of CSR with fixed n

data:  airbnb0619SG_ppp
R = 0.36169, p-value = 0.01
alternative hypothesis: clustered (R < 1)

##### 5.2.1.2.2 Clark and Evans Test: Kallang planning area

```{r}
clarkevans.test(airbnb0619_kl_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: 

##### 5.2.1.2.3 Clark and Evans Test: River Valley planning area

```{r}
clarkevans.test(airbnb0619_rv_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: 

##### 5.2.1.2.4 Clark and Evans Test: Novena planning area

```{r}
clarkevans.test(airbnb0619_nv_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: 

##### 5.2.1.2.5 Clark and Evans Test: Rochor planning area

```{r}
clarkevans.test(airbnb0619_rc_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: 


### G-FUNCTION

#### Hypothesis

#### Testing and Analysis

### F-FUNCTION

#### Hypothesis

#### Testing and Analysis

### K-FUNCTION

#### Hypothesis

#### Testing and Analysis

### L-FUNCTION

#### Hypothesis

#### Testing and Analysis








TBC... NOT SURE
### 5.1.2 Kernel density maps analysis on Openstreetmap of Singapore to see possible relationship

```{r}
tmap_mode("plot")
```

Airbnb0619 and Hotel

```{r}
airbnb0619map <- tm_shape(kde_airbnb0619SG_bw_raster) +
                  tm_raster("v", palette="Purples", colorNA = NULL) +
                  tm_basemap("OpenStreetMap")


hotelsmap <- tm_shape(kde_hotelsSG_bw_raster) +
                  tm_raster("v", palette="Purples", colorNA = NULL, showNA = NA) +
                  tm_basemap("OpenStreetMap") 

tmap_arrange(airbnb0619map, hotelsmap)
```

Airbnb0619 and MrtLrt

```{r}
airbnb0619map <- tm_shape(kde_airbnb0619SG_bw_raster) +
                  tm_raster("v", palette="Purples", colorNA = NULL) +
                  tm_basemap("OpenStreetMap")


mrtlrtmap <- tm_shape(kde_mrtlrtSG_bw_raster) +
                  tm_raster("v", palette="Purples", colorNA = NULL, showNA = NA) +
                  tm_basemap("OpenStreetMap") 

tmap_arrange(airbnb0619map, mrtlrtmap)
```

Airbnb0619 and Tourist Attractions

```{r}
airbnb0619map <- tm_shape(kde_airbnb0619SG_bw_raster) +
                  tm_raster("v", palette="Purples", colorNA = NULL) +
                  tm_basemap("OpenStreetMap")


tourismmap <- tm_shape(kde_tourismSG_bw_raster) +
                  tm_raster("v", palette="Purples", colorNA = NULL, showNA = NA) +
                  tm_basemap("OpenStreetMap") 

tmap_arrange(airbnb0619map, tourismmap)
```

```{r}
tmap_mode("plot")
```
