---
title: "Hands-on_Ex08"
description: |
  I have learnt how to perform geographical segmentation, cluster analysis and visualising clustering results.
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 10-10-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Installing Packages

```{r}
packages = c('rgdal', 'spdep', 'tmap', 'sf', 'ggpubr', 'cluster', 'factoextra', 'NbClust', 'heatmaply', 'corrplot', 'psych', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
    }
  library(p,character.only = T)
}
```

# Data Preparation

## Geospatial

```{r}
shan_sf <- st_read(dsn = "data/geospatial", layer = "myanmar_township_boundaries") %>%
  filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)"))
```

```{r}
shan_sf
```

```{r}
glimpse(shan_sf)
```

## Aspatial

```{r}
ict <- read_csv ("data/aspatial/Shan-ICT.csv")
```

```{r}
summary(ict)
```

New variable with dplyr (decimal scaling)

```{r}
ict_derived <- ict %>%
  mutate(`RADIO_PR` = `Radio`/`Total households`*1000) %>%
  mutate(`TV_PR` = `Television`/`Total households`*1000) %>%
  mutate(`LLPHONE_PR` = `Land line phone`/`Total households`*1000) %>%
  mutate(`MPHONE_PR` = `Mobile phone`/`Total households`*1000) %>%
  mutate(`COMPUTER_PR` = `Computer`/`Total households`*1000) %>%
  mutate(`INTERNET_PR` = `Internet at home`/`Total households`*1000) %>%
  rename(`DT_PCODE` =`District Pcode`,`DT`=`District Name`,
         `TS_PCODE`=`Township Pcode`, `TS`=`Township Name`,
         `TT_HOUSEHOLDS`=`Total households`,
         `RADIO`=`Radio`, `TV`=`Television`, 
         `LLPHONE`=`Land line phone`, `MPHONE`=`Mobile phone`,
         `COMPUTER`=`Computer`, `INTERNET`=`Internet at home`) 
```

```{r}
summary(ict_derived)
```

# EDA

## Statistical Graph

Histogram (radio)

```{r}
ggplot(data=ict_derived, aes(x=`RADIO`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

Box Plot (radio)

```{r}
ggplot(data=ict_derived, aes(x=`RADIO`)) +
  geom_boxplot(color="black", fill="light blue")
```

Histogram (radio_pr)

```{r}
ggplot(data=ict_derived, aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

Box Plot (radio)

```{r}
ggplot(data=ict_derived, aes(x=`RADIO_PR`)) +
  geom_boxplot(color="black", fill="light blue")
```

Multiple histogram (to see data range)

```{r}
radio <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

tv <- ggplot(data=ict_derived, 
             aes(x= `TV_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

llphone <- ggplot(data=ict_derived, 
             aes(x= `LLPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

mphone <- ggplot(data=ict_derived, 
             aes(x= `MPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

computer <- ggplot(data=ict_derived, 
             aes(x= `COMPUTER_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

internet <- ggplot(data=ict_derived, 
             aes(x= `INTERNET_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```


```{r}
ggarrange(radio, tv, llphone, mphone, computer, internet, 
          ncol = 3, 
          nrow = 2)
```

## Choropleth map

Join Geospatial with Aspatial

```{r}
shan_sf <- left_join(shan_sf, ict_derived, by=c("TS_PCODE"="TS_PCODE"))
```

Choropleth map (Radio_PR)

```{r}
qtm(shan_sf, "RADIO_PR")
```

Choropleth side by side (Total Households, Number Radio)

```{r}
TT_HOUSEHOLDS.map <- tm_shape(shan_sf) + 
  tm_fill(col = "TT_HOUSEHOLDS",
          n = 5,
          style = "jenks", 
          title = "Total households") + 
  tm_borders(alpha = 0.5) 

RADIO.map <- tm_shape(shan_sf) + 
  tm_fill(col = "RADIO",
          n = 5,
          style = "jenks",
          title = "Number Radio ") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(TT_HOUSEHOLDS.map, RADIO.map,
             asp=NA, ncol=2)
```

Choropleth side by side (Total Households, Radio_PR)

```{r}
tm_shape(shan_sf) +
    tm_polygons(c("TT_HOUSEHOLDS", "RADIO_PR"),
                style="jenks") +
    tm_facets(sync = TRUE, ncol = 2) +
  tm_legend(legend.position = c("right", "bottom"))+
  tm_layout(outer.margins=0, asp=0)
```

# Correlation Analysis

```{r}
cluster_vars.cor = cor(ict_derived[,12:17])
corrplot.mixed(cluster_vars.cor,
         lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

# Hierarchy Cluster Analysis

## Extract clustering variables

```{r}
cluster_vars <- shan_sf %>%
  st_set_geometry(NULL) %>%
  select("TS.x", "RADIO_PR", "TV_PR", "LLPHONE_PR", "MPHONE_PR", "COMPUTER_PR")
head(cluster_vars,10)
```

Change rows by township name

```{r}
row.names(cluster_vars) <- cluster_vars$"TS.x"
head(cluster_vars,10)
```

delete the TS.x field 

```{r}
shan_ict <- select(cluster_vars, c(2:6))
head(shan_ict, 10)
```

# Data Standardization

## Min-Max standardisation

```{r}
shan_ict.std <- normalize(shan_ict)
summary(shan_ict.std)
```

## Z-score standardisation

```{r}
shan_ict.z <- scale(shan_ict)
describe(shan_ict.z)
```

## Visualising the standardised clustering variables

default vs min-max vs z-score

```{r}
r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)
```

# Computing proximity matrix

euclidean method

```{r}
proxmat <- dist(shan_ict, method = 'euclidean')
proxmat
```

# Computing hierarchical clustering

Hierarchical cluster analysis using ward.D
 
```{r}
hclust_ward <- hclust(proxmat, method = 'ward.D')
```

```{r}
plot(hclust_ward, cex = 0.6)
```

# Selecting the optimal clustering algorithm

Compute the agglomerative coefficients of all hierarchical clustering algorithms.

```{r}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

ac <- function(x) {
  agnes(shan_ict, method = x)$ac
}

map_dbl(m, ac)
```

# Determining Optimal Clusters

Gap Statistic Method

Compute gap statistic

```{r}
set.seed(12345)
gap_stat <- clusGap(shan_ict, FUN = hcut, nstart = 25, K.max = 10, B = 50)
# Print the result
print(gap_stat, method = "firstmax")
```

```{r}
fviz_gap_stat(gap_stat)
```

# Interpreting the dendrograms

draw the dendrogram with border around the selected clusters 

Note that width can change but height cannot...

```{r}
plot(hclust_ward, cex = 0.6)
rect.hclust(hclust_ward, k = 6, border = 2:5)
```

# Visually-driven hierarchical clustering analysis

## Transforming the data frame into a matrix

```{r}
shan_ict_mat <- data.matrix(shan_ict)
```

## Plotting interactive cluster heatmap using heatmaply()

```{r}
heatmaply(normalize(shan_ict_mat),
          Colv=NA,
          dist_method = "euclidean",
          hclust_method = "ward.D",
          seriate = "OLO",
          colors = Blues,
          k_row = 6,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="Geographic Segmentation of Shan State by ICT indicators",
          xlab = "ICT Indicators",
          ylab = "Townships of Shan State"
          )
```

# Mapping the clusters formed

derive 5-cluster model

```{r}
groups <- as.factor(cutree(hclust_ward, k=6))
```

```{r}
shan_sf_cluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER`=`as.matrix.groups.`)
```

```{r}
qtm(shan_sf_cluster, "CLUSTER")
```

# Spatially Constrained Clustering - SKATER approach

## Converting into SpatialPolygonsDataFrame

```{r}
shan_sp <- as_Spatial(shan_sf)
```

## Converting into SpatialPolygonsDataFrame

```{r}
shan.nb <- poly2nb(shan_sp)
summary(shan.nb)
```

```{r}
plot(shan_sp, border=grey(.5))
plot(shan.nb, coordinates(shan_sp), col="blue", add=TRUE)
```

## Computing minimum spanning tree

### Calculating edge costs

Compute cost of each edge

```{r}
lcosts <- nbcosts(shan.nb, shan_ict)
```

Incorporate costs into weight objects

```{r}
shan.w <- nb2listw(shan.nb, lcosts, style="B")
summary(shan.w)
```

### Computing minimum spanning tree

Computing with mean of mstree()

```{r}
shan.mst <- mstree(shan.w)
```

Check its class and dimension

```{r}
class(shan.mst)
```

```{r}
dim(shan.mst)
```

display the content of shan.mst

```{r}
head(shan.mst)
```

```{r}
plot(shan_sp, border=gray(.5))
plot.mst(shan.mst, coordinates(shan_sp), 
     col="blue", cex.lab=0.7, cex.circles=0.005, add=TRUE)
```

# Computing spatially constrained clusters using SKATER method

compute the spatially constrained cluster using skater()

```{r}
clust6 <- skater(shan.mst[,1:2], shan_ict, method = "euclidean", 5)
```

```{r}
str(clust6)
```

check the cluster assignment 

```{r}
ccs6 <- clust6$groups
ccs6
```

observations are in each cluster

```{r}
table(ccs6)
```

plot pruned tree of the five clusters

```{r}
plot(shan_sp, border=gray(.5))
plot(clust6, coordinates(shan_sp), cex.lab=.7,
     groups.colors=c("red","green","blue", "brown", "pink"), cex.circles=0.005, add=TRUE)
```

# Visualising the clusters in choropleth map

```{r}
groups_mat <- as.matrix(clust6$groups)
shan_sf_spatialcluster <- cbind(shan_sf_cluster, as.factor(groups_mat)) %>%
  rename(`SP_CLUSTER`=`as.factor.groups_mat.`)
qtm(shan_sf_spatialcluster, "SP_CLUSTER")
```
Side by side comparison 

```{r}
hclust.map <- qtm(shan_sf_cluster,
                  "CLUSTER") + 
  tm_borders(alpha = 0.5) 

shclust.map <- qtm(shan_sf_spatialcluster,
                   "SP_CLUSTER") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(hclust.map, shclust.map,
             asp=NA, ncol=2)
```
