---
title: "In Class Exercise 5"
description: |
  A short description of the post.
author:
  - name: Kwek Yi Chen
    url: https://example.com/kwekyichen
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE, eval=TRUE, echo=TRUE, message=FALSE, error=FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = FALSE)
```


## Installing and Loading the R package

+ maptools: for creating spatial* objects
+ sf: importing geospatial data
+ raster: creating raster object
+ spatstat: performing Spatial point pattern analysis
+ tmap: creating/plotting thematic maps
+ tidyverse: data handling
+ plotly: plot interactive plot
+ ggthemes: advance theme for plot

```{r}
packages <- c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse', 'plotly', 'ggthemes')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

## Import the Geospatial data

Importing shapefile using *st_read()* of **sf** package. The output object is in tibble sf object class.

```{r}
mpsz_sf <- st_read(dsn="data/shapefile",
                     layer="MP14_SUBZONE_WEB_PL")
```

Projection is in svy21.

## Importing th Aspatial data from rds folder

*read_rds()* of **readr** package is used instead of *readRDS()* of base R. Ouput of *read_rds()* is in tibble object.

```{r}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```

Note that there are some data issue in childcare data frame because `Lat` and `Lng` should be in numeric data type. The coordinate fields seems to be in decimal degree. Hence, wgs referencing system is assumed.

## Converting the aspatial data frame into sf object

```{r}
CHAS_sf <- st_as_sf(CHAS,
                    coords = c("X_COORDINATE", 
                               "Y_COORDINATE"),
                    crs=3414)
```

Note: *st_as_sf()* accept coordinates in character data type

```{r}
childcare_sf <- st_as_sf(childcare,
                    coords = c("Lng", 
                               "Lat"),
                    crs=4326) %>%
  st_transform(crs = 3414)
```


## Plotting to review the chilcare_sf

alpha sets transparency.

```{r}
tmap_mode("view")
tm_shape(childcare_sf)+
  tm_dots(alpha=0.4,
          col = "blue",
          size = 0.05) +
tm_shape(CHAS_sf)+
  tm_dots(alpha=0.4,
          col = "red",
          size = 0.05)
```

## Geospatial Data Wrangling


### Converting sf to Spatial* data frame 

*as_Spatial()* of sf package

```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

chas: spatialpointsdf
chilcare: spatialpointsdf
mpsz: spatialpolygondf

### Converting Spatial* df into Spaital* objects

*as.SpatialPoint()* or *as.SpatialPolygon()* of maptools package

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

### Converting from Spatial* object to ppp object

using as.ppp() of maptools package

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```

### Removing duplicate points using jitter

*rjitter()* of spatstat function

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp,
                            retry = TRUE,
                            nsim = 1,
                            drop = TRUE)

any(duplicated(childcare_ppp_jit))
```

```{r}
CHAS_ppp_jit <- rjitter(CHAS_ppp,
                            retry = TRUE,
                            nsim = 1,
                            drop = TRUE)

any(duplicated(CHAS_ppp_jit))
```

tmap dont understand ppp...

### Extracting Punggol Planning Area

, is necessary to complete syntax

```{r}
pg <- mpsz[mpsz@data$PLN_AREA_N=="PUNGGOL",]
```

### Converting SpatialPolygonDataFrame into SpatialPolygons Object

```{r}
pg_sp <- as(pg, "SpatialPolygons")
```

### Converting SpatialPolygons into owin object

owin() of spatstat. need owin to define study area

```{r}
pg_owin <- as(pg_sp, "owin")
```

### Extracting spatial points within owin

```{r}
childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]
```

```{r}
plot(childcare_pg)
```

### L-funtion

```{r}
L_childcare <- envelope(childcare_pg,
                        Lest,
                        nsim=99,
                        rank=1,
                        global = TRUE)

L_CHAS <- envelope(CHAS_pg,
                        Lest,
                        nsim=99,
                        rank=1,
                        global = TRUE)

```

